### 2024-12-03 paper/paper.qmd
results section completed

---
title: "Barriers to Green Living: Insights from 2018 and 2021 Toronto Climate Perception Surveys"
subtitle: "Uncovering What Stops Sustainable Behaviors and How Communities Can Overcome Them"
author: 
  - Lexi Knight
thanks: "Code and data are available at: [https://github.com/LexiKnight/toronto_climate/tree/main](https://github.com/LexiKnight/toronto_climate/tree/main)."
date: today
date-format: long
abstract: "This paper explores the factors limiting individuals from adopting environmentally friendly behaviors, drawing on survey data to understand motivations and barriers. Findings reveal that while many people are aware of climate change, obstacles like cost, convenience, and perceived effectiveness hinder sustainable actions. Additionally, the research highlights which methods of communication are most effective across different age groups. By addressing these barriers and tailoring strategies for engagement, cities can better empower residents to contribute to meaningful climate action, fostering a more sustainable urban future."
format:
  pdf:
    toc: true
number-sections: true
bibliography: references.bib
---

# Introduction

Climate change remains one of the most pressing global challenges, with rising temperatures, extreme weather events, and environmental degradation threatening ecosystems and human well-being. Despite the urgent need for individual and collective action, many people continue to struggle with adopting sustainable behaviors in daily life. Governments and communities are increasingly interested in understanding how to motivate environmentally friendly practices, but significant gaps remain in identifying and addressing the obstacles that prevent action. This paper seeks to bridge that gap by examining not only what prevents individuals from engaging in sustainable behaviors but also what communication strategies can effectively inspire change.

This study focuses on a survey conducted in 2018, which gathered insights into people’s awareness of climate change, their likelihood of taking specific climate-friendly actions, the reasons behind inaction, preferred ways of receiving climate-related information, and demographic characteristics such as age and education. We investigate the extent to which people feel informed about climate change, their readiness to adopt energy-efficient practices, reduce waste, and embrace greener transportation options, among other behaviors. Our analysis also sheds light on the barriers individuals face, ranging from financial and logistical concerns to doubts about the impact of their actions, and explores how demographic factors influence communication preferences.

Addressing the gap between climate awareness and action is critical for effective community climate initiatives. Studies show that while awareness of environmental issues is widespread, behavioral change remains limited (Gifford, 2011). Additionally, research indicates that communication strategies tailored to specific demographics can enhance engagement and effectiveness (Moser & Dilling, 2007). Yet, there is little empirical evidence that explores how personalized communication and targeted interventions can overcome barriers to sustainable behaviors. Our study aims to fill this gap by estimating the causal impact of various factors on individuals’ willingness and ability to engage in environmentally friendly practices and identifying the best strategies for promoting sustainable change.

We found that practical obstacles, such as cost and convenience, alongside perceptions of ineffectiveness, are the primary reasons for inaction. Moreover, age and education level significantly shape how people prefer to receive information about climate action, suggesting that a one-size-fits-all communication strategy may be insufficient. These findings have substantial implications for policymakers and community leaders seeking to design more impactful climate initiatives. By addressing key barriers and tailoring engagement strategies, cities can empower individuals to adopt sustainable behaviors, making meaningful progress toward climate resilience and environmental sustainability.

In Canada, electric vehicles (EVs) are increasingly recognized as a more sustainable alternative to internal combustion engine (ICE) vehicles. A life cycle assessment (LCA) conducted by the National Research Council of Canada highlights that EVs offer substantial greenhouse gas (GHG) emission reductions, particularly when powered by electricity from low-carbon sources, such as hydroelectricity in provinces like Québec. The study emphasizes the importance of considering regional electricity mixes, which significantly influence the environmental performance of EVs compared to conventional vehicles (NRC, 2021)
Moreover, advancements in battery technology are further enhancing the sustainability of EVs. New lithium iron phosphate (LFP) batteries, which reduce the need for critical metals like cobalt and nickel, are becoming more prevalent. These batteries not only lower production costs but also extend the lifespan of EVs, making them even more environmentally friendly over time. (NRC, 2021)
In Ontario, where Toronto is located, the electricity grid's relatively low carbon intensity makes EVs a cleaner option throughout most of their lifecycle. According to Clean Energy Canada, EVs in the country can emit up to 89% fewer GHGs than gasoline-powered cars, even after accounting for emissions from battery production and vehicle manufacturing (NRC, 2021)

The remainder of this paper is structured as follows: @sec-data presents the methods used to analyze these demographic differences, followed by @sec-results, which details the results of the comparison between Rosedale-Moore Park and Downtown Yonge East. @sec-discussion offers a discussion of the findings, highlighting the implications for understanding residential segregation in Toronto. Finally, @sec-conclusion concludes with a reflection on the limitations of the study and suggestions for future research directions.

# Data {#sec-data}
The primary data source for this project is the Climate Perception Study dataset, provided by the City of Toronto via the Open Data Toronto platform [@climateperceptionstudy]. This dataset offers insights into public perceptions of climate change in Toronto and was accessed from Open Data Toronto [@opendatatoronto] on 6 November 2024.
## Software and R-packages
This project was conducted using the statistical software `R` [@citeR]. Data cleaning and manipulation were primarily facilitated by the `tidyverse` package [@tidyverse], which encompasses various essential tools. Within this suite, `dplyr` [@dplyr] was employed for data manipulation tasks such as filtering, summarizing, and joining datasets, while `readr` [@readr] provided efficient functionality for reading and writing rectangular text data. Specifically, the `write_csv` function from `tidyverse` was used to save the raw 2018 dataset. To handle Excel files, we incorporated the `httr` package for downloading files directly from online sources and `readxl` [@readxl], which enabled us to read Excel files seamlessly. Additionally, the `openxlsx` package was utilized to create a new Excel workbook with multiple sheets, allowing us to store the 2021 dataset in a structured format.

The `arrow` package was employed to handle large datasets efficiently, ensuring smooth data processing, while `forcats` was used for reordering factor levels to facilitate more insightful analyses and visualizations. In terms of data tidying, the `tidyr` and `stringr` packages allowed us to reshape and clean the data by managing character strings and handling missing values. For the final documentation and report generation, the `knitr` package was instrumental in integrating R code and outputs into a cohesive and reproducible report format. This suite of R packages collectively enabled us to conduct thorough data management, manipulation, and analysis, ultimately contributing to the successful completion of the project.

## Methodology 
### Data Collection
The data for this analysis comes from two surveys commissioned by the City of Toronto to assess residents' perceptions of climate change and their willingness to take community action. In 2018, Environics Research conducted n=404 online interviews with Toronto residents aged 18 and older between October 11 and October 18. Each interview lasted approximately 10 minutes. The sample was drawn from an online panel, with quotas set by region, age, and gender to reflect the city’s population based on the 2016 Census, and minor weighting was applied. Due to the non-random nature of the online survey, no margin of error was assigned, though a random sample of this size would typically have a margin of ±4.9 percentage points, 19 times out of 20.

In 2021, a second survey was conducted by Ipsos, a global market research company. This survey included n=1,400 Toronto residents aged 18 and older and used a mixed-method approach. The sample consisted of 1,000 respondents from an online panel, 300 recruited via cell phone and landline, and 100 online interviews conducted in Mandarin, Cantonese, and Punjabi. This combination of methods aimed to enhance the survey's representativeness. For analysis, Version 2 of the 2021 dataset, which presents responses in a numeric format, was used.

### Data Cleaning
#### 2018 Individual Data
The cleaning process for the 2018 individual data began with selecting six key questions from the survey, focusing on variables that would provide insights into individual perceptions and behaviors regarding climate change. The corresponding columns were extracted and renamed with meaningful, descriptive names for clarity. The selected variables included age, the extent to which individuals consider themselves informed about the causes of climate change, the likelihood of taking specific climate actions, reasons for inaction (if they indicated they were unlikely to act), the highest level of education completed, and preferred methods for the city to deliver information about climate change and climate action.

After selecting and renaming the columns, minor formatting inconsistencies were addressed. For instance, a missing space in the "verylikely" response category was corrected to "very likely." Additionally, certain values were reformatted to ensure compatibility with visualization tools, making the dataset easier to plot later. Unlike other datasets that often contain missing values or duplicate entries, this dataset was notably clean. There were no missing responses, and all questions were answered, allowing for a comprehensive analysis of survey participants' perceptions and behaviors. The dataset offered valuable insights into individuals’ likelihood of engaging in climate actions and their preferred communication channels for climate information.

Throughout the cleaning process, no observations were removed, as every response was preserved to accurately represent the survey data. Despite encountering a few incorrect variable formats, these were easily corrected without data loss. The dataset provided a rare opportunity to analyze a complete and consistent set of individual responses, enabling a thorough exploration of both demographic characteristics and personal perceptions related to climate change.\

#### 2018 Summary Data
Cleaning Process for the 2018 Summary Data
To create the 2018 summary data, I built off the cleaned 2018 individual data. This approach was necessary because the 2021 data was only available in summarized form, not at the individual level. By summarizing the 2018 data in a comparable format, I could facilitate a meaningful comparison between the two years. The first step involved loading the cleaned 2018 individual dataset and mimicking the structure of the 2021 summary data as closely as possible.

For age, I created the same age categories used in the 2021 dataset to ensure consistency between the two datasets. I then calculated the percentage of individuals within each age group and created a table summarizing this information. Similarly, I summarized the highest level of education completed by respondents, presenting it as a percentage of the total. However, a slight inconsistency emerged during this process, as the education levels and their descriptions differed between 2018 and 2021. Despite this discrepancy, I aligned the categories as closely as possible to maintain comparability.

Next, I calculated the percentage of respondents who reported feeling informed about the causes of climate change and summarized the likelihood of taking various climate actions. This step was somewhat more complex, as the dataset contained nine different actions, each rated on a five-point scale of likelihood. I created a table showing the percentage likelihood of taking each specific action. Additionally, I analyzed the reasons respondents provided for being unlikely to take certain actions, which were only collected if they had indicated “unlikely” in the previous question. These reasons were summarized as counts rather than percentages, reflecting the conditional nature of the responses.

Finally, I summarized the preferred methods for receiving climate-related information from the city, presenting this data as a percentage. This comprehensive summarization of the 2018 individual data provided a consistent and comparable dataset for analyzing trends and changes in perceptions and behaviors between 2018 and 2021.

#### 2021 Summary Data 
The cleaning process for the 2021 summary data focuses on ensuring consistency and accuracy for six specific variables: age, education, extent informed, likelihood to take action, and reasons for not taking action. I first extracted data from the relevant sheets and columns in the 2021 raw dataset using read_excel from the tidyverse. For each category, I selected specific rows and converted raw data into percentages. The age and education data were cleaned and saved into separate data frames (age_summary_21 and education_summary_21), while other categories like "extent informed" and "likelihood to take action" required cleaning across multiple columns, with data transformation into percentages. Missing values and outliers were handled by filtering invalid entries and ensuring all values were consistent.

Instead of merging datasets for 2018 and 2021, I aligned the 2018 summary data to match the structure and categories of the 2021 data to facilitate comparison. This alignment was necessary because the categories and structure of the two datasets differed, and I modified the 2018 dataset to resemble the 2021 format as closely as possible, allowing for a more direct comparison. By focusing on ensuring that categories and percentages were comparable between years, I preserved the integrity of both datasets.

Finally, I saved each cleaned dataset into formatted tables using tinytable for LaTeX compatibility. These cleaned datasets were essential for analysis, ensuring that all variables were represented as percentages without any missing or erroneous data. This process also ensured that the dataset was ready for comparison with the 2018 data, with clear, consistent categories across both years.

### Data Analysis
The primary focus of this analysis was the 2018 dataset, examining six key variables: age, education, extent informed, likelihood to take action, reasons for not taking action, and the best method for delivering information regarding climate change action. I analyzed these variables at the individual level in the 2018 dataset by calculating the percentages within each category, which provided a clear understanding of the distribution and trends within the population.

In addition, I compared the 2018 dataset to the 2021 summary data to assess changes over the years. By calculating the percentage distribution for each variable in both years, I was able to identify any shifts in public awareness, engagement, and preferences regarding climate change action. This comparison revealed whether there have been notable changes in factors like the public’s level of information about climate change, their likelihood to take action, and their preferred communication methods.

This two-step approach allowed for a comprehensive understanding: a detailed exploration of the 2018 data combined with a broader comparison of trends between 2018 and 2021.

## Features
### Age
#### Individual 2018
@fig-one illustrates the age distribution of survey respondents in 2018. The histogram displays the frequency of respondents across different age groups, with the x-axis representing age intervals and the y-axis indicating the number of respondents within each interval. The distribution is relatively balanced, with two noticeable peaks around the 30–40 and 50–60 age ranges, suggesting these age groups are the most represented in the survey.

The vertical dotted lines in the figure highlight key statistical measures: the green dotted line represents the median age, while the blue dashed line indicates the mean age of the respondents. Both measures fall near the center of the distribution, reinforcing the observation that the sample is primarily composed of middle-aged individuals. Although the distribution appears fairly symmetrical, there is a gradual decline in respondent frequency after the age of 60, with fewer participants aged 90 and above, indicating a smaller representation of older individuals.

Overall, @fig-one provides valuable context for understanding the demographic composition of the survey sample. The predominance of middle-aged respondents may influence the interpretation of survey results, particularly if age is a factor in attitudes, behaviors, or preferences related to the study.

```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-one
#| fig-cap: The histogram shows the distribution of age given the 2018 survey respondents.

knitr::include_graphics(here("data/03-figures_data", "age_individual_plot.png"))

```
#### Comparison 2018 to 2021
@table-one illustrates the different age ditribution for 2018 and 2021 surveys. The major difference is that there was greater representation of younger inviduals ranging from 18 to 23 in 2021 compared to 2018. The remaining age groups were not very different.

```{r}
#| warning: false
#| echo: false
#| message: false
#| label: table-one
print(age_summary_combined)

```

### Highest Level of Education Completed
#### Individual 2018
@fig-two illustrates the distribution of survey respondents by their highest level of education attained. The most frequently reported education levels are "Completed undergraduate degree" and "Post graduate/professional school," with both categories having a noticeably higher frequency compared to others, indicating a well-educated respondent population. In contrast, fewer respondents reported "Some community college, vocational, trade school" and "Completed community college, vocational, trade school," suggesting these levels are less common among participants. Additionally, a small portion of respondents preferred not to disclose their education level. This distribution highlights a skew toward higher educational attainment within the sample, which may influence responses related to socio-economic factors and other survey topics.
```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-two
#| fig-cap: The histogram shows the distribution of education levels of 2018 survey respondents.

knitr::include_graphics(here("data/03-figures_data", "educ_individual_plot.png"))

```
#### Comparison 2018 to 2021
@table-two illustrates respondant education level in 2018 and 2021 respectively. Please note, the categories of education level are different for 2018 and 2021 as these were different surveys and categories were updated in 2021. Despite this, we are still able to make comparisons. Individuals with high school diplomas or less account for 9 for 2018 and 38 for 2021 (addition of less than highs school and high school or equivalent categories). This could indicate that the younger population was included more in the 2021 suvey or that less education individuals were surveyed. Additionally, its not super clear howver those with a post graduate or professional degree accounted for 29 percent in 2018 and 35 percent in 2021. This illustrates that more educated inviduals responded to the 2021 survey.
```{r}
#| warning: false
#| echo: false
#| message: false
#| label: table-two
#| fig-cap:?

print(education_summary_combined)
```

### Extent Feel Informed
#### Individual 2018
@fig-three depicts the distribution of self-reported climate change knowledge among survey respondents in 2018. The majority of participants identified as "Very informed," making this the most prevalent category by a significant margin. A moderate number of respondents reported being "Not very informed," while a smaller yet noticeable group described themselves as "Extremely informed." In contrast, the categories "Not at all informed" and "Not very informed" have the lowest frequencies, with "Not at all informed" being particularly rare. This distribution suggests that most respondents possess a relatively high level of awareness about climate change, with very few indicating a lack of knowledge on the topic.
```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-three
#| fig-cap: The histogram shows the distribution of self-reported climate change knowledge among 2018 survey respondents.

knitr::include_graphics(here("data/03-figures_data", "informed_individual_plot.png"))

```
#### Comparison 2018 to 2021
@table-three illustrates the self-reported level of extent inviduals feel informed about climate change and climate action for 2018 and 2021. In 2018, the large marjority of individuals say they are very informed with 60 percent and the follwoign category is not very informed with 27 percent. In 2021, The majority of inividuals also say they are very informed but to a lesser degree with 48 percent and this is closely followed by only a little informed with 37 percent. There is a trend that invidiusls in 2018 reported greater levels of knowldege about climate change whereas 2021 illstrates a more even split between very informed and only little informed, portraying that fewer individuals have adequate knowledge about climate change.

```{r}
#| warning: false
#| echo: false
#| message: false
#| label: table-three
#| fig-cap:?

print(informed_summary_combined)
```

### Best Mode of Delivery
#### Individual 2018
The analysis of the best method for delivering climate change information portrayed in @fig-four revealed that the majority of individuals in the 2018 dataset preferred receiving information firstly from advertising campaigns with about 55 percent, newsletters with about 52 percent, followed by the toronto website wiht about 46 percent. On the other side of the spectrum, few said they were uninterested in receiving climate action information  with about 5 percent, about 12 percent indicated twitter and then 17 percent mentioned instagram. Overall social media platforms Facebook, Instagram and Twtitter were less popular methods of communciation. This could be because the survey only included individals over the age of 18. Moreover, the median sample was in the 40s.

```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-four
#| fig-cap: The stacked bar plot illustrates the preferred methods of communication for receiving information about climate change and climate action.

knitr::include_graphics(here("data/03-figures_data", "communication_18_plot.png"))

```

#### Comparison 2018 to 2021
@table-four compares responses from 2018 and 2021 surverys regarding the best method of delivering information on climate change and climate action. Aside from the additon of four options to the 2021 survey, both surveys questions are the same. The leading category for 2018 is advertising campaigns with 55 percent, closely followed by newsletter emails with 52 percent and the Toronto website with 46. The 2021 survey illustrates a consistent trend with the same platforms as the top method of delivering information, specifically the Toronto website with 44 percent, then enewsletter emails with 43 percent and advertising campains at 42 percent. These top methods of communciation are more evenly spread out than the 2018 survey results. Social media platforms namely twitter, facebook and instagram all increased in the 2021 survey. Notable decreases in term of percentage from 2018 to 2021 include events, enewsletter emails, advertising campaigns, brochures and pamphlets. The new category of BetterHomesTO.ca webiste accounts for 15 percent of votes for the 2021 survey.

```{r}
#| warning: false
#| echo: false
#| message: false
#| label: table-four
#| fig-cap:?

print(communication_summary_combined_table)
```
### Likelihood Taking Action
#### Individual 2018
The likelihood of taking action against climate change was explored by assessing individuals' intentions and behaviors. 
@fig-five illustrates self-reported likelihood of taking certain actions that minimize climate change effects. On the x-axis, we have the nine different climate actions and on the y-axis is the percentage. The stacked bar chart allows understanding of how likely individuals are to take part in the given actions. 
Sorting waste into the correct bins is the action that most individuals are doing right now, accounting for 60 percent whith 24 percent saying they are very likely to take part in this action. The next most prominent actions are walking or cycling short distances, reduce hydro usage, reduce waste and home improvements. 40 percent of invidiuals report walking or cycling when traveling shorter distances in the city with 22 percent reporting they are very likely. However, 7 percent say they are very unlikely to change their method of transportation. Reducing the amount of energy and water use is already done by 36 percent and very likely for 29 percent only a mere 3 percent is very unlikely. Reducing personal waste such as contributing to repurposing used products has similar percentages to responses for reducing hydro but even fewer individuals that report they are very unlikely to take action. Similarly, the likelihoods for making home improvements for energy efficiency such as energy-efficient appliances, programmable thermostat, LED lightbulbs, green or cool roof. 
In contrast, the action where the least number of individuals are contributing to is the investment of an electric or hybrid vehicle over gas-powered vehicles with only 8 percent who currently have an energy efficient vehicle. While 19 percetn are very likely to make this switch, 14 percent say they are very unlikely. This illustrates that buying an electric or hybrid vehicle is action the least number of individuals are willing to invest in despite gas-powered vehicles being a major contributor to greenhouse gases. This action has a huge impact on the environment and yet 41 percent of this sample say they are unlikely to change! The second action that individuals are most unlikely to engage in is eat less meat by incorporating more plant-based foods in their diet with 12 percent very unlikely and 23 percent somewhat unlikely. Only 23 percent are already consuming meat alternatives. Vegans greenhosue gas emissions are on average 37% lower than those who eat meat frequently. 
```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-five
#| fig-cap: The stacked bar plot shows the likelihood of survey respondents taking various actions to address climate change.

knitr::include_graphics(here("data/03-figures_data", "likelihood_individual_plot.png"))

```

# Model {#sec-model}
TO DO:
The goal of our modelling strategy is twofold. Firstly,...
Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-appendix].
## Model set-up
In this analysis, we aim to predict the likelihood of taking climate-friendly actions based on demographic variables (age, education) and climate change awareness (knowledge). The specific actions include investing in an electric or hybrid vehicle instead of a gas-powered one, making home improvements (e.g., using LED lightbulbs), adopting meat alternatives, minimizing car use, purchasing green products, reducing hydro usage, reducing waste by repurposing used products, sorting waste correctly, and walking or cycling shorter distances.

### Model Specifications ###
TO DO:

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model Justification 
This model is designed to explore how various demographic and knowledge-based factors influence the likelihood of engaging in climate-friendly behaviors. The likelihood of taking specific actions, such as reducing energy consumption or investing in green technologies, represents an important outcome in understanding climate change perceptions. As such, focusing on these behaviors rather than general knowledge allows for a more actionable and directly applicable model to climate change policy and interventions.

#### Response Variable ####
The response variable in this analysis is the likelihood of taking climate-friendly actions, which is typically a continuous or ordinal measure based on respondents' self-reported likelihood of adopting behaviors such as switching to an electric vehicle, reducing waste, or conserving energy. These actions are commonly measured on a Likert scale (e.g., 1 = "not likely" to 5 = "very likely").

#### Input Variabless ####
The input variables (predictors) include:

Age: Previous studies indicate that younger individuals are more likely to engage in climate-positive behaviors due to greater environmental awareness and concern (Leiserowitz et al., 2010; Gifford, 2013).

Education: Higher educational attainment has been associated with a higher likelihood of adopting environmentally responsible behaviors, likely due to increased awareness of environmental issues and access to relevant information (Kollmuss & Agyeman, 2002).

Knowledge: A greater understanding of climate change is expected to positively influence individuals' likelihood of taking action. More knowledgeable individuals are more likely to adopt climate-friendly practices, as they are more aware of the urgency of the crisis and the potential benefits of sustainable behaviors (Gifford, 2013).

#### Model Structure ####
The model assumes that the likelihood of taking climate-friendly actions is linearly related to the input variables. Each predictor (age, education, and knowledge) is assumed to have a linear relationship with the likelihood of taking action, meaning that as these variables change, the likelihood of taking climate-friendly actions will either increase or decrease in a predictable manner. This structure facilitates a clear interpretation of how each variable contributes to the overall likelihood of climate-positive behavior.

Detailed Structure
The model does not assume any interactions between the predictors, meaning each factor (age, education, knowledge) is considered independently in its influence on the outcome. However, depending on the data, you might consider adding interaction terms if you believe that, for example, the effect of knowledge on action is stronger for younger individuals.
The linear relationship between predictors and the likelihood of action allows for easy interpretation, with coefficients representing the change in the likelihood of taking action for each unit change in the predictor variable.

#### Parameter Estimation ####
In parameter estimation, we anticipate that:
Age: We expect that younger individuals will have a higher likelihood of taking climate-friendly actions, as studies have shown that younger generations are more environmentally conscious and concerned about climate change (Leiserowitz et al., 2010).
Education: We hypothesize that higher levels of education will positively influence the likelihood of taking climate-positive actions. More educated individuals tend to have better access to information about climate change and sustainability practices (Kollmuss & Agyeman, 2002).
Knowledge: We expect that greater knowledge of climate change will be strongly associated with higher likelihoods of adopting climate-friendly behaviors. Knowledge acts as a critical enabler, motivating individuals to take action and contribute to mitigating climate change (Gifford, 2013).
We will estimate the model parameters using a standard linear regression approach (e.g., using the lm() function in R), which will provide the coefficients (
β
β) for each predictor, showing how each factor influences the likelihood of taking climate-friendly actions.

# Results

## Likelihood taking action by age
@fig-six shows the relationship between age and the likelihood of engaging in various climate-related actions. The x-axis represents the age groups, and the y-axis indicates the percentage of respondents reporting different likelihood levels, from "Already doing this" to "Very unlikely." The trend lines across each action show how engagement varies by age.

### Electric / Hybrid Vehicle Adoption
Younger respondents (15-24, 25-34) are less likely to adopt electric or hybrid vehicles, with only a small proportion already using or considering them. Adoption rates increase slightly in the 55-64 age group, then decline again in the 65+ group. This suggests that older individuals may have more resources or stability to invest in electric vehicles.

### Home Improvement for Energy Efficiency
Middle-aged individuals (35-54) are the most likely to engage in home improvement actions aimed at increasing energy efficiency. Younger age groups (15-24, 25-34) report lower likelihoods, likely due to barriers such as homeownership and disposable income.

### Consumption of Meat Alternatives
Younger age groups, especially 15-24 and 25-34, are more likely to adopt meat alternatives. As age increases, the likelihood of doing so decreases, possibly due to established dietary habits in older generations.

### Minimizing Car Use
A trend similar to that of meat alternatives is observed, with younger individuals more likely to minimize car use. Older respondents, particularly those 65 and older, show less inclination to reduce car usage, possibly due to mobility needs or lifestyle factors like suburban living.

### Purchasing Green Products
The likelihood of purchasing green products is relatively uniform across age groups. Middle-aged and older demographics show a slight increase in engagement, but the trend line remains mostly flat, indicating that this behavior is more influenced by factors such as income or environmental awareness rather than age.

### Reducing Hydro Usage
Older respondents (55-64 and 65+) are the most likely to reduce hydro usage, likely driven by cost-saving measures or environmental concerns. The trend line shows a steady increase with age.

### Reducing and Sorting Waste
Both reducing and sorting waste are more commonly reported by middle-aged and older groups (45-64). This may reflect increased awareness or resources for waste management.

### Walking / Cycling Short Distances
 Younger respondents (15-34) are more likely to engage in walking or cycling short distances. As age increases, the likelihood of walking or cycling decreases, possibly due to physical limitations or differing lifestyle needs.

```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-six
#| fig-cap: Likelihood of Taking Climate Change Actions by Age Group. The stacked bar chart illustrates the percentage of respondents across six age groups (15-24, 25-34, 35-44, 45-54, 55-64, 65+) in terms of their self-reported likelihood to engage in nine distinct climate-related actions. Each color represents a different level of likelihood, from "Already doing this or have done this" to "Very unlikely." The black dashed trend line highlights how the likelihood of action changes across age groups, providing insights into generational differences in climate-related behavior.

knitr::include_graphics(here("data/03-figures_data", "likelihood_age_plot.png"))
```

## Likelihood taking action by education 
@fig-seven displays how the likelihood of engaging in climate-friendly actions varies by educational attainment. The x-axis represents the proportion of respondents in each education category, while the y-axis lists the educational levels, from "Prefer not to answer" to "Postgrad." Each action is shown across separate facets, revealing how education level correlates with engagement in each behavior.

### Trends in Education Attainment and Climate-Friendly Actions
Individuals with higher education levels generally show a higher likelihood of engaging in climate-friendly actions. This trend is evident across all behaviors, with respondents possessing undergraduate or postgraduate degrees more likely to adopt electric vehicles or purchase green products.

#### Electric/Hybrid Vehicle Adoption
Postgraduate degree holders are the most likely to have adopted or consider adopting electric or hybrid vehicles. In contrast, respondents with lower education levels, such as those with high school or less, are more likely to report being "Very unlikely" or "Somewhat unlikely" to adopt these vehicles.

#### Minimizing Car Use
Higher education levels correlate with a greater likelihood of minimizing car use, though the trend is less pronounced for individuals with lower education levels.

#### Reducing Hydro Usage
While education has a smaller impact on reducing hydro usage compared to other actions, those with higher education levels are still more likely to engage in this behavior.

#### Purchasing Green Products and Reducing Waste
Purchasing green products and reducing waste are more common among individuals with postsecondary education, especially those with undergraduate or postgraduate degrees. In contrast, individuals with lower education levels report more moderate engagement with these actions.

### Adopting Meat Alternatives
Higher education is associated with greater adoption of meat alternatives. This is particularly true for those with undergraduate or postgraduate education, while fewer respondents with lower education levels report engaging in this behavior.

### Walking or Cycling Short Distances
Walking or cycling short distances is a behavior that is relatively common across all education levels, though individuals with higher education levels show slightly higher engagement.

### Sorting Waste Correctly
Sorting waste correctly is one of the most consistently adopted behaviors across all education levels, with a relatively small proportion of respondents indicating they are unlikely to engage in this action. However, individuals with higher education are still more likely to fall in the "Already doing this" category.

```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-seven
#| print(likelihood_education_plot)
```

## Overall Summary of Results:
The results indicate that both age and education significantly influence the likelihood of engaging in climate-related actions. Younger respondents tend to be more likely to adopt behaviors such as consuming meat alternatives and minimizing car use, while older individuals are more inclined to engage in actions that require larger investments or long-term commitments, such as reducing hydro usage and investing in energy-efficient home improvements. Higher educational attainment is also strongly linked to a greater likelihood of engaging in climate-friendly behaviors, particularly those that involve financial investment, like adopting electric vehicles and consuming meat alternatives. However, simpler actions such as sorting waste correctly are more widely reported across all education levels. These generational and educational differences highlight the importance of factors such as income, lifestyle, and environmental awareness in shaping individuals' climate-related behaviors. In light of these findings, it is crucial to consider demographic factors when designing policies or campaigns aimed at encouraging climate action, as different age and education groups exhibit varied levels of engagement with both simple and more complex sustainable practices.

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Limitations

no individual data for 2021, had to used summary statistics instead
i was originally very interested in loooking at the reasons behind why people are unlikely to perform such climate friendly actions

## Future Directions

# Conclusion {#sec-conclusion}

\newpage

\appendix

# Appendix {#sec-appendix}

\newpage

# References