### 2024-12-11 paper/climate_change.qmd
asjust cross referencing

---
title: "Predicting the Likelihood of Climate-Friendly Actions: Insights from Age and Educational Attainment"
subtitle: "An analysis of how demographic factors influence the adoption of sustainable behaviors"
author: 
  - Lexi Knight
thanks: "Code and data are available at: [https://github.com/LexiKnight/toronto_climate/tree/main](https://github.com/LexiKnight/toronto_climate/tree/main)."
date: today
date-format: long
abstract: "This study examines the relationship between demographic factors—age, education, and self-reported knowledge of climate change—and individuals' likelihood of engaging in climate-friendly behaviors. The analysis reveals that younger people and those with higher levels of education are more likely to adopt sustainable actions such as reducing car use and purchasing electric vehicles. However, the study also highlights that financial barriers and limited access to resources remain significant obstacles for broader climate action. These findings underscore the importance of tailoring policies and interventions to make sustainable behaviors more accessible to diverse demographic groups, promoting wider participation in efforts to combat climate change."
format:
  pdf:
    toc: true
number-sections: true
bibliography: references.bib
---

# Introduction

Climate change is one of the most urgent challenges we face today. From rising temperatures to extreme weather events, the planet is already feeling the consequences of human-driven environmental change [@IPCC2023]. Yet, despite widespread awareness of these risks, many people still struggle to adopt behaviors that could help mitigate climate damage [@Gifford2011]. This gap between knowledge and action is a central puzzle that governments, communities, and organizations are trying to solve [@Kollmuss2002]. If we are to tackle this crisis effectively, we need to better understand what stops people from acting, and how we can motivate change [@Steg2014].

This paper examines how specific factors—such as age, education, and self-reported knowledge of climate change—shape people’s willingness to engage in climate-friendly actions [@Schultz2002]. By analyzing data from a 2018 survey [@climateperceptionstudy], we explore the connection between demographic characteristics and behaviors like reducing energy consumption, using greener transportation, and lowering waste. The survey also delves into what prevents people from adopting these behaviors, uncovering obstacles like cost, convenience, and skepticism about the power of individual actions [@Gifford2011]. With this analysis, we aim to uncover how these factors interact and what they mean for efforts to encourage sustainable choices [@Stern2000].

The heart of this study is to measure how age, education, and knowledge of climate change influence the likelihood of adopting certain behaviors. We found that younger individuals and those with higher education levels are more likely to engage in actions like reducing car usage and adopting plant-based diets [@Fulton2022]. However, older individuals are more inclined to take steps like cutting down on household energy use [@Lee2019]. Beyond these demographic trends, we also discovered that barriers such as financial cost and doubts about the effectiveness of individual efforts often hold people back from taking action [@Kollmuss2002]. These findings suggest that a “one-size-fits-all” approach to climate communication won’t work. Instead, targeted strategies that consider demographic differences are essential for driving meaningful engagement [@Steg2014].

Understanding these patterns is crucial for crafting effective climate policies. By identifying which groups are more likely to act and which face significant barriers, we can better design interventions that speak to people’s concerns and motivations [@Stern2000]. The implications of this study go beyond simply knowing what people think about climate change—they offer a roadmap for shaping messages and policies that could shift behavior on a larger scale [@Fulton2022].

The remainder of this paper is structured as follows: @sec-data details the data and methods used in our analysis. @sec-model describes how we assess the factors influencing climate-friendly behaviors across demographic groups. @sec-results highlights key differences in behavior based on age, education, and knowledge of climate change. @sec-discussion discusses the findings, addresses the study's limitations, and suggests directions for future research. @sec-conclusion concludes with a summary of the key takeaways and actionable recommendations for enhancing climate action engagement. Finally, @sec-appendix contains supplementary material.

# Data {#sec-data}
The primary data source for this project is the Climate Perception Study dataset, provided by the City of Toronto via the Open Data Toronto platform [@climateperceptionstudy]. This dataset offers insights into public perceptions of climate change in Toronto and was accessed from Open Data Toronto [@opendatatoronto] on 6 November 2024.
## Software and R-packages
This project was conducted using the statistical software `R` [@citeR]. Data cleaning and manipulation were primarily facilitated by the `tidyverse` package [@tidyverse], which encompasses various essential tools. Within this suite, `dplyr` [@dplyr] was employed for data manipulation tasks such as filtering, summarizing, and joining datasets, while `readr` [@readr] provided efficient functionality for reading and writing rectangular text data. Specifically, the `write_csv` function from `tidyverse` was used to save the raw 2018 dataset. To handle Excel files, we incorporated the `httr` package for downloading files directly from online sources and `readxl` [@readxl], which enabled us to read Excel files seamlessly. Additionally, the `openxlsx` package was utilized to create a new Excel workbook with multiple sheets, allowing us to store the 2021 dataset in a structured format.

The `arrow` package was employed to handle large datasets efficiently, ensuring smooth data processing, while `forcats` was used for reordering factor levels to facilitate more insightful analyses and visualizations. In terms of data tidying, the `tidyr` and `stringr` packages allowed us to reshape and clean the data by managing character strings and handling missing values. The [@rpart2024] and [@partykit2024] were used in to make the models. For the final documentation and report generation, the `knitr` package was instrumental in integrating R code and outputs into a cohesive and reproducible report format. This suite of R packages collectively enabled us to conduct thorough data management, manipulation, and analysis, ultimately contributing to the successful completion of the project.

## Methodology 
### Data Collection
The data for this analysis comes from two surveys commissioned by the City of Toronto to assess residents' perceptions of climate change and their willingness to take community action. In 2018, Environics Research conducted n=404 online interviews with Toronto residents aged 18 and older between October 11 and October 18. Each interview lasted approximately 10 minutes. The sample was drawn from an online panel, with quotas set by region, age, and gender to reflect the city’s population based on the 2016 Census, and minor weighting was applied. Due to the non-random nature of the online survey, no margin of error was assigned, though a random sample of this size would typically have a margin of ±4.9 percentage points, 19 times out of 20.

In 2021, a second survey was conducted by Ipsos, a global market research company. This survey included n=1,400 Toronto residents aged 18 and older and used a mixed-method approach. The sample consisted of 1,000 respondents from an online panel, 300 recruited via cell phone and landline, and 100 online interviews conducted in Mandarin, Cantonese, and Punjabi. This combination of methods aimed to enhance the survey's representativeness. For analysis, Version 2 of the 2021 dataset, which presents responses in a numeric format, was used.

### Data Analysis
The primary focus of this analysis was the 2018 dataset, examining six key variables: age, education, extent informed, likelihood to take action, reasons for not taking action, and the best method for delivering information regarding climate change action. I analyzed these variables at the individual level in the 2018 dataset by calculating the percentages within each category, which provided a clear understanding of the distribution and trends within the population.

In addition, I compared the 2018 dataset to the 2021 summary data to assess changes over the years. By calculating the percentage distribution for each variable in both years, I was able to identify any shifts in public awareness, engagement, and preferences regarding climate change action. This comparison revealed whether there have been notable changes in factors like the public’s level of information about climate change, their likelihood to take action, and their preferred communication methods.

This two-step approach allowed for a comprehensive understanding: a detailed exploration of the 2018 data combined with a broader comparison of trends between 2018 and 2021.

## Features
### Age
#### Individual 2018
@fig-one illustrates the age distribution of survey respondents in 2018. The histogram displays the frequency of respondents across different age groups, with the x-axis representing age intervals and the y-axis indicating the number of respondents within each interval. The distribution is relatively balanced, with two noticeable peaks around the 30–40 and 50–60 age ranges, suggesting these age groups are the most represented in the survey.

The vertical dotted lines in the figure highlight key statistical measures: the green dotted line represents the median age, while the blue dashed line indicates the mean age of the respondents. Both measures fall near the center of the distribution, reinforcing the observation that the sample is primarily composed of middle-aged individuals. Although the distribution appears fairly symmetrical, there is a gradual decline in respondent frequency after the age of 60, with fewer participants aged 90 and above, indicating a smaller representation of older individuals.

Overall, @fig-one provides valuable context for understanding the demographic composition of the survey sample. The predominance of middle-aged respondents may influence the interpretation of survey results, particularly if age is a factor in attitudes, behaviors, or preferences related to the study.

```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-one
#| fig-cap: "The histogram shows the distribution of age given the 2018 survey respondents."

library(here)

knitr::include_graphics(here("data/03-figures_data", "age_individual_plot.png"))

```
\newpage

#### Comparison 2018 to 2021
@fig-two illustrates the different age ditribution for 2018 and 2021 surveys. The major difference is that there was greater representation of younger inviduals ranging from 18 to 23 in 2021 compared to 2018. The remaining age groups were not very different.

```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-two
#| fig-cap: "This table highlights the age distribution of survey respondents in 2018 and 2021."

# load libraries 
library(kableExtra)
library(knitr)
library(here)
library(dplyr)

# Read the cleaned CSV into a data frame
age_summary_combined <- read.csv(here("data/03-figures_data", "age_summary_combined.csv"))

# Rename columns to "Age Group", "2018 Percentage", and "2021 Percentage"
colnames(age_summary_combined) <- c("Age Group", "2018 Percentage", "2021 Percentage")

# Print the cleaned table (using kable for styling)
age_summary_combined %>%
  kbl() %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))

```

### Highest Level of Education Completed
#### Individual 2018
@fig-three illustrates the distribution of survey respondents by their highest level of education attained. The most frequently reported education levels are "Completed undergraduate degree" and "Post graduate/professional school," with both categories having a noticeably higher frequency compared to others, indicating a well-educated respondent population. In contrast, fewer respondents reported "Some community college, vocational, trade school" and "Completed community college, vocational, trade school," suggesting these levels are less common among participants. Additionally, a small portion of respondents preferred not to disclose their education level. This distribution highlights a skew toward higher educational attainment within the sample, which may influence responses related to socio-economic factors and other survey topics.
```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-three
#| fig-cap: "The histogram shows the distribution of education levels of 2018 survey respondents."

knitr::include_graphics(here("data/03-figures_data", "educ_individual_plot.png"))

```
#### Comparison 2018 to 2021
@fig-four illustrates respondant education level in 2018 and 2021 respectively. Please note, the categories of education level are different for 2018 and 2021 as these were different surveys and categories were updated in 2021. Despite this, we are still able to make comparisons. Individuals with high school diplomas or less account for 9 for 2018 and 38 for 2021 (addition of less than highs school and high school or equivalent categories). This could indicate that the younger population was included more in the 2021 suvey or that less education individuals were surveyed. Additionally, its not super clear howver those with a post graduate or professional degree accounted for 29 percent in 2018 and 35 percent in 2021. This illustrates that more educated inviduals responded to the 2021 survey.
```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-four
#| fig-cap: This table highlights the education distribution of survey respondents in 2018 and 2021."

# Load necessary libraries
library(tinytable)

# Read the CSV into a data frame
education_summary_combined <- read.csv(here("data/03-figures_data", "education_summary_combined.csv"))

# Rename columns 
colnames(education_summary_combined) <- c("Education Level ", "2018 Percentage", "Education Level", "2021 Percentage")

# Convert the data frame to tinytable
education_summary_combined_tiny <- tt(education_summary_combined)

# Print the tinytable
education_summary_combined_tiny

```

### Extent Feel Informed
#### Individual 2018
@fig-five depicts the distribution of self-reported climate change knowledge among survey respondents in 2018. The majority of participants identified as "Very informed," making this the most prevalent category by a significant margin. A moderate number of respondents reported being "Not very informed," while a smaller yet noticeable group described themselves as "Extremely informed." In contrast, the categories "Not at all informed" and "Not very informed" have the lowest frequencies, with "Not at all informed" being particularly rare. This distribution suggests that most respondents possess a relatively high level of awareness about climate change, with very few indicating a lack of knowledge on the topic.
```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-five
#| fig-cap: The histogram shows the distribution of self-reported climate change knowledge among 2018 survey respondents.

knitr::include_graphics(here("data/03-figures_data", "informed_individual_plot.png"))

```
#### Comparison 2018 to 2021
@fig-six illustrates the self-reported level of extent inviduals feel informed about climate change and climate action for 2018 and 2021. In 2018, the large marjority of individuals say they are very informed with 60 percent and the follwoign category is not very informed with 27 percent. In 2021, The majority of inividuals also say they are very informed but to a lesser degree with 48 percent and this is closely followed by only a little informed with 37 percent. There is a trend that invidiusls in 2018 reported greater levels of knowldege about climate change whereas 2021 illstrates a more even split between very informed and only little informed, portraying that fewer individuals have adequate knowledge about climate change.

```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-six
#| fig-cap: "This table highlights the extent feel informed distribution of survey respondents in 2018 and 2021."

# Load libraries
library(kableExtra)
library(knitr)
library(here)
library(dplyr)    # For renaming columns
library(arrow)    # For reading Parquet files

# Read the Parquet file into a data frame
informed_summary_combined <- arrow::read_parquet(here("data/03-figures_data", "informed_summary_combined_table.parquet"))

# Rename columns
colnames(informed_summary_combined) <- c("Extent Informed", "2018 Percentage", "Extent Informed", "2021 Percentage")

# Print the table (using kable for styling)
informed_summary_combined %>%
  kbl() %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))

```

### Best Mode of Delivery
#### Individual 2018
T@ fig-seven, shows the analysis of the best method for delivering climate change information portrayed in @fig-four revealed that the majority of individuals in the 2018 dataset preferred receiving information firstly from advertising campaigns with about 55 percent, newsletters with about 52 percent, followed by the toronto website wiht about 46 percent. On the other side of the spectrum, few said they were uninterested in receiving climate action information  with about 5 percent, about 12 percent indicated twitter and then 17 percent mentioned instagram. Overall social media platforms Facebook, Instagram and Twtitter were less popular methods of communciation. This could be because the survey only included individals over the age of 18. Moreover, the median sample was in the 40s.

```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-seven
#| fig-cap: The stacked bar plot illustrates the preferred methods of communication for receiving information about climate change and climate action.

knitr::include_graphics(here("data/03-figures_data", "communication_18_plot.png"))

```

#### Comparison 2018 to 2021
@fig-eight compares responses from 2018 and 2021 surverys regarding the best method of delivering information on climate change and climate action. Aside from the additon of four options to the 2021 survey, both surveys questions are the same. The leading category for 2018 is advertising campaigns with 55 percent, closely followed by newsletter emails with 52 percent and the Toronto website with 46. The 2021 survey illustrates a consistent trend with the same platforms as the top method of delivering information, specifically the Toronto website with 44 percent, then enewsletter emails with 43 percent and advertising campains at 42 percent. These top methods of communciation are more evenly spread out than the 2018 survey results. Social media platforms namely twitter, facebook and instagram all increased in the 2021 survey. Notable decreases in term of percentage from 2018 to 2021 include events, enewsletter emails, advertising campaigns, brochures and pamphlets. The new category of BetterHomesTO.ca webiste accounts for 15 percent of votes for the 2021 survey.

```{r}
#| warning: false
#| echo: false
#| message: false
#| label: table-eight
#| fig-cap: "This table highlights the prefered method of communication of survey respondents in 2018 and 2021."

# Load libraries
library(kableExtra)
library(knitr)
library(here)
library(dplyr)    # For renaming columns
library(arrow)    # For reading Parquet files

# Read the Parquet file into a data frame
communication_summary_combined <- arrow::read_parquet(here("data/03-figures_data", "communication_summary_combined_table.parquet"))

# Rename columns
colnames(communication_summary_combined) <- c("Communication Method", "2018 Percentage", "Communication Method", "2021 Percentage")

# Print the table (using kable for styling)
communication_summary_combined %>%
  kbl() %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))

```
### Likelihood Taking Action
#### Individual 2018
The likelihood of taking action against climate change was explored by assessing individuals' intentions and behaviors. 
@fig-nine illustrates self-reported likelihood of taking certain actions that minimize climate change effects. On the x-axis, we have the nine different climate actions and on the y-axis is the percentage. The stacked bar chart allows understanding of how likely individuals are to take part in the given actions. 
Sorting waste into the correct bins is the action that most individuals are doing right now, accounting for 60 percent whith 24 percent saying they are very likely to take part in this action. The next most prominent actions are walking or cycling short distances, reduce hydro usage, reduce waste and home improvements. 40 percent of invidiuals report walking or cycling when traveling shorter distances in the city with 22 percent reporting they are very likely. However, 7 percent say they are very unlikely to change their method of transportation. Reducing the amount of energy and water use is already done by 36 percent and very likely for 29 percent only a mere 3 percent is very unlikely. Reducing personal waste such as contributing to repurposing used products has similar percentages to responses for reducing hydro but even fewer individuals that report they are very unlikely to take action. Similarly, the likelihoods for making home improvements for energy efficiency such as energy-efficient appliances, programmable thermostat, LED lightbulbs, green or cool roof. 
In contrast, the action where the least number of individuals are contributing to is the investment of an electric or hybrid vehicle over gas-powered vehicles with only 8 percent who currently have an energy efficient vehicle. While 19 percetn are very likely to make this switch, 14 percent say they are very unlikely. This illustrates that buying an electric or hybrid vehicle is action the least number of individuals are willing to invest in despite gas-powered vehicles being a major contributor to greenhouse gases. This action has a huge impact on the environment and yet 41 percent of this sample say they are unlikely to change! The second action that individuals are most unlikely to engage in is eat less meat by incorporating more plant-based foods in their diet with 12 percent very unlikely and 23 percent somewhat unlikely. Only 23 percent are already consuming meat alternatives. Vegans greenhosue gas emissions are on average 37% lower than those who eat meat frequently. 
```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-nine
#| fig-cap: The stacked bar plot shows the likelihood of survey respondents taking various actions to address climate change.

knitr::include_graphics(here("data/03-figures_data", "likelihood_individual_plot.png"))

```

# Model {#sec-model}
The goal of our modelling strategy is twofold. Firstly, we aim to predict the likelihood of individuals engaging in climate-friendly actions based on demographic variables, specifically age and education. The actions considered include investing in electric or hybrid vehicles, making home improvements (such as using LED lightbulbs), adopting meat alternatives, minimizing car use, purchasing green products, reducing hydro usage, reducing waste, sorting waste correctly, and walking or cycling shorter distances. Secondly, we aim to understand how these demographic characteristics influence individuals' likelihood of adopting these behaviors.Here we describe the probability tree model used to investigate the relationship between likelihood and the demographic factors of age and education.

## Model set-up
In this analysis, we focus on predicting the likelihood of taking climate-friendly actions based on the demographic variables of age and education. The likelihood is measured on a five-point Likert scale, ranging from "Very unlikely" to "Already doing this or have done this."

### Model Specifications ###
We use a probability tree model to predict the likelihood of individuals engaging in climate-friendly actions based on two key demographic variables: age and education.

The general form of the model is:

\begin{align} 
y_i = \beta_0 + \beta_1 * Age_i + \beta_2 * Education_i + \epsilon_i
\end{align*}
where:

 - $y_i$ represents the likelihood of individual $i$ engaging in a climate-friendly actin, measured on a five point Likert scale:
 1. "Very unlikely"
 2. "Somewhat unlikely"
 3. "Somewhat likely"
 4. "Very likely"
 5. "Already doing this or have done this"
- $\beta_0$ is the intercept, representing the baseline likelihood of taking action when all predictors are zero.
- $\beta_1$ is the coefficient for Age, representing how a one-year increase affects the likelihood of taking action.
- $\beta_2$ is the coefficient for Education, a categorical variable that is dummy-coded.
- $\epsilon_i$ is the error term, accounting for unexplained variation.

This model focuses on the direct effect of age and education on the likelihood of taking climate-friendly actions, assuming a linear relationship between these demographic variables and the outcome variable (the likelihood of action).

### Model Justification 
This model was developed to explore how demographic factors such as age and education influence individuals' likelihood of engaging in climate-friendly behaviors. By focusing on specific climate actions like reducing energy consumption, adopting sustainable practices, and investing in green technologies, the model aims to provide actionable insights for climate change interventions.

#### Response Variable ####
The response variable in this analysis is the likelihood of taking climate-friendly actions, measured using a Likert scale (1 = "Very unlikely" to 5 = "Already doing this or have done this"). This scale captures the respondent’s perceived likelihood of engaging in various climate-friendly behaviors

#### Input Variabless ####
The input variables (predictors) include:

- **Age**: Previous research indicates that younger individuals may be more inclined to adopt climate-friendly behaviors due to increased environmental awareness (Leiserowitz et al., 2010; Gifford, 2013).
- **Education**: Higher levels of education are often linked to greater awareness of environmental issues and a higher likelihood of adopting sustainable behaviors (Kollmuss & Agyeman, 2002). Education is treated as a categorical variable and dummy-coded for analysis.

#### Model Structure ####
This model assumes a linear relationship between the predictors (age and education) and the likelihood of taking action. Specifically, the likelihood of taking climate-friendly actions is modeled as a function of age and education, with coefficients representing the change in the likelihood of action associated with a unit change in each predictor.

There are no interaction terms in this model, meaning we assume that the effects of age and education on the likelihood of taking action are independent of each other. This simplicity ensures clear interpretation of each predictor's individual effect on the likelihood of taking climate-friendly actions.

## Nine Models ##

We fit nine separate probability tree models, one for each climate-friendly action we are predicting. Each model uses age and education as the predictors to estimate the likelihood of the corresponding action. These actions are:

1. Likelihood of investing in an electric or hybrid vehicle
2. Likelihood of making home improvements (e.g., using LED lightbulbs)
3. Likelihood of adopting meat alternatives
4. Likelihood of minimizing car use
5. Likelihood of purchasing green products
6  Likelihood of reducing hydro usage
7. Likelihood of reducing waste by repurposing used products
8. Likelihood of sorting waste correctly
9. Likelihood of walking or cycling shorter distances

Each model is built using the same structure but predicts a different outcome variable. This approach allows us to analyze the specific impact of age and education on each climate-friendly action.

### Decision Trees
A decision tree is a machine learning model used to make predictions by recursively splitting data into subsets based on specific conditions of the predictor variables. It is structured like a tree, where the top node is the root node, representing the entire dataset, and subsequent nodes represent splits based on decision rules. The branches indicate the outcomes of these rules, and the leaf nodes (terminal nodes) provide the final prediction for each subset.

Each split is determined by the predictor variable most strongly associated with the target outcome, maximizing predictive accuracy. A split separates individuals by age or educational level. The prediction in each terminal node includes the majority class (e.g., "Very likely") and additional information such as sample size (n) and error percentage. The error percentage reflects the proportion of cases in that node misclassified by the model. A higher error percentage indicates greater uncertainty or variability within the group, while a lower error percentage suggests a more precise classification.

#### Likelihood of investing in an electric or hybrid vehicle
@fig-ten predicts individuals’ likelihood to invest in an electric or hybrid vehicle based on their highest level of education and age. The model identifies highest level of education as the most important variable for the initial split. At the root node, individuals are divided into two groups: those with "Some Community/Trade School" or "Some University" education and those with other educational backgrounds such as "High School," "Undergrad," "Postgrad," or "Prefer Not to Answer."

For individuals with "Some Community/Trade School" or "Some University" education, the tree predicts that they are "Very likely" to invest, with a sample size of 47 and an error rate of 63.8%. This indicates moderate variability within this group.

For those with other educational backgrounds, further splits are based on age. Younger individuals (below 38.5 years) are segmented further. A subset aged under 34.5 years is split into two groups, with one group (n = 10) being "Very likely" to invest, with a lower error rate of 30%. The other group is classified as "Somewhat likely," with a higher error rate of 60.7%. Older individuals (38.5 years or older) are less likely to invest. For instance, those aged above 68.5 years are predicted as "Very unlikely" with a sample size of 18 and an error rate of 61.1%, while those aged between 34.5 and 68.5 years are divided into "Somewhat unlikely" or "Somewhat likely" categories.

Overall, the decision tree reveals distinct patterns where educational attainment and age interact to predict investment likelihood, with younger individuals and those with less formal education being more likely to invest.

```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-ten
#| fig-cap: Probability tree for likelihood of investing in an electric or hybrid vehicle, predicted by age and education.

knitr::include_graphics(here::here("data/04-model_data/vehicle_electric_tree.png"))

```

#### Likelihood of making home improvements
In @fig-eleven, the root node splits the data at age 35.5, indicating that age is the most influential factor in determining climate-related home improvement behavior. Individuals aged 36 years or older are predominantly classified as "Already doing this or have done this," with a large sample size (n = 223) but a moderate error rate of 59.6%, reflecting variability in this group.

For individuals younger than 36 years, the tree introduces education as the next important factor. Those with higher educational attainment ("Community/Trade School," "Some University," or "Postgraduate") are classified as "Very likely" to make climate-conscious home improvements (n = 40) with an error rate of 55.0%. Conversely, individuals with lower education levels, such as "High School" or "Some Community/Trade School," are classified as "Somewhat likely" (n = 46) but have a higher error rate of 67.4%, indicating less predictive certainty. A smaller subset of younger individuals with similar educational attainment is classified as "Somewhat likely" (n = 10) with a lower error rate of 40.0%, suggesting higher confidence in this classification. The decision tree highlights clear trends based on age and education, with older age groups and higher educational attainment linked to greater engagement in making home improvements.

```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-eleven
#| fig-cap: Probability tree for likelihood of making home improvements predicted by age and education.

knitr::include_graphics(here::here("data/04-model_data/home_improvement_tree.png"))

```

#### Likelihood of adopting meat alternatives
@fig-twelve predicts individuals’ likelihood of consuming meat alternatives to minimize the effects of climate change based on their age and highest level of education. The model identifies age as the most important variable for the initial split. At the root node, individuals are divided into two groups: those under 34.5 years old and those 34.5 years or older. This highlights age as a critical factor influencing the likelihood of adopting meat alternatives.

In the subgroup of individuals under 34.5 years old, the decision tree further splits based on age at 31.5 years, classifying younger individuals into categories of “Somewhat likely” (n = 25, error = 44%) and “Somewhat unlikely” (n = 62, error = 62.9%). On the other hand, for individuals aged 34.5 years or older, the model considers highest level of education as the next key variable, separating those with community/trade school, undergraduate, and postgraduate education from those with high school education. This distinction reflects how educational attainment moderates the impact of age on the decision to adopt meat alternatives.

Within the older age group, additional splits occur based on age, such as at 49 years and 65.5 years, leading to further classifications like “Already doing this or have done this” or “Somewhat likely.” These splits indicate how the interaction between age and education level shapes patterns of meat alternative adoption, resulting in more granular predictions.

```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-twelve
#| fig-cap: Probability tree for likelihood of adopting meat alternatives, predicted by age and education.

knitr::include_graphics(here::here("data/04-model_data/protein_alternative_tree.png"))

```

#### Likelihood of minimizing car use
@fig-thirteen predicts individuals' likelihood of minimizing car use as a measure to mitigate climate change. In this case, the tree structure is straightforward, containing only a single root node. The model classifies all individuals into one category, "Already doing this or have done this," with a sample size of 320 and an error rate of 61.9%. The absence of splits in the tree indicates that no other variables, such as age or education, contributed sufficient predictive power to justify further segmentation.

This outcome suggests that either the input data lacks variability in relevant features, or the behavior of minimizing car use is widespread and consistent across the sample population. Consequently, the model does not differentiate between subgroups and treats all individuals as having the same likelihood of engaging in this behavior.
```{r}
#| warning: false
#| echo: false
#| message: false
#| label: @fig-thirteen
#| fig-cap: Probability tree for likelihood of minimizing car use, predicted by age and education.

knitr::include_graphics(here::here("data/04-model_data/minimize_car_tree.png"))

```

#### Likelihood of purchasing green products
@fig-fourteen predicts individuals' likelihood of purchasing green products to minimize the effects of climate change based on age and highest level of education. The model identifies age as the most important variable for the initial split. At the root node, individuals are divided into two groups: those younger than 37.5 years old and those 37.5 years or older. This division highlights age as a key determinant in predicting green product purchasing likelihood.

For individuals aged 37.5 years or older, the next split is determined by the highest level of education. Those with educational backgrounds such as "High School," "Postgrad," "Some Community/Trade School," "Some University," or "Prefer Not to Answer" are further analyzed. Within this group, additional splits are made based on age thresholds of 50.5 years and 47.5 years. These splits refine the classification into categories such as "Somewhat likely" or "Already doing this or have done this," with varying error rates and sample sizes.

Among individuals younger than 37.5 years, the highest level of education again plays a crucial role in the split. Individuals with educational backgrounds such as "Postgrad," "Community/Trade School," "High School," or "Some University" are further segmented. Within this subgroup, age thresholds of 27.5 years and 22.5 years are used to differentiate individuals into categories such as "Very likely" or "Somewhat likely," reflecting diverse levels of engagement with green product purchasing. Overall, the decision tree incorporates age and education as the primary variables to segment individuals into likelihood categories.
```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-forteen
#| fig-cap: Probability tree for likelihood of purchasing green products, predicted by age and education.

knitr::include_graphics(here::here("data/04-model_data/green_product_tree.png"))

```

#### Likelihood of reducing hydro usage
@fig-fifteen identifies age and highest level of education as key predictors influencing the likelihood of reducing hydro usage. The root node of the tree reveals that individuals younger than 34.5 years old are categorized as "Very likely" to reduce hydro usage, with a sample size of 87 and an error rate of 60.9%. This suggests that younger individuals may demonstrate a greater propensity to adopt hydro-reduction practices. Conversely, individuals over 34.5 years are further split by age, with those older than 64.5 years also categorized as "Very likely," albeit with a smaller sample size of 16 and a lower error rate of 37.5%. This bifurcation highlights that age groups at the extremes of the spectrum (younger and older individuals) are more inclined toward hydro-use reduction.

Among middle-aged individuals (34.5–64.5 years), the decision tree highlights the role of education. Those younger than 61.5 years and possessing higher levels of education, including university degrees, are either "Already doing this or have done this" (n=151, err=56.3%) or "Somewhat likely" (n=14, err=50.0%) to engage in hydro-use reduction. On the other hand, individuals with lower educational attainment, particularly trade school or community-level education, are less likely to implement such measures. For individuals aged 61.5–71.5 years, there remains a balance between those already reducing hydro usage and those who are very likely to adopt it in the future, with clear variability based on educational level.
```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-fifteen
#| fig-cap: Probability tree for likelihood of reducing hydro usage, predicted by age and education.

knitr::include_graphics(here::here("data/04-model_data/reduce_hydro_tree.png"))

```

####  Likelihood of reducing waste
@fig-sixteen predicts the likelihood of reducing waste is based on age and highest level of education. The root node splits individuals by age at 34.5 years, indicating that younger and older individuals exhibit distinct waste-reduction behaviors. For those under 34.5 years, the model uses highest level of education as a key determinant, classifying them as either "Somewhat likely" or "Very likely" to reduce waste.

For individuals aged 34.5 years or older, the tree splits again by education and further refines predictions using additional age thresholds at 61.5 and 75.5 years. Older individuals with higher education levels are more likely to have "Already done this or are doing this," while those with lower education levels are categorized as "Somewhat likely" or "Very likely." This structure highlights the combined influence of education and age on waste-reduction behaviors, with clear thresholds that differentiate groups.

The decision tree demonstrates that age is the primary driver of waste-reduction behavior, with education acting as an important secondary factor. These variables together capture differences in sustainability habits, showing how younger, middle-aged, and older populations vary in their likelihood to engage in reducing waste.
```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-sixteen
#| fig-cap: Probability tree for likelihood of reducing waste, predicted by age and education.

knitr::include_graphics(here::here("data/04-model_data/reduce_waste_tree.png"))

```

####  Likelihood of sorting waste correctly
@fig-seventeen predicts the likelihood of sorting waste correctly identifies highest level of education as the primary determinant. Individuals with education levels at or above Community/Trade School, Postgraduate, or Some University are categorized as "Already doing this or have done this," accounting for the majority of respondents (n = 295, err = 37.3%). This demonstrates that higher educational attainment correlates with proactive waste-sorting behaviors.

For individuals whose highest level of education is High School, the model introduces a secondary split based on age at 37 years. Younger individuals under 37 years are categorized as "Already doing this or have done this," albeit with a higher error rate (n = 7, err = 57.1%). Older individuals aged 37 or above are more likely classified as "Very likely" to sort waste correctly (n = 17, err = 35.3%). These splits highlight the combined influence of education and age on waste-sorting behavior, with specific subgroups showing varying levels of commitment and accuracy.

This decision tree underscores that education plays a dominant role in predicting waste-sorting behavior, while age further refines these predictions, especially for those with lower education levels. The model provides straightforward insights into how demographics shape sustainable practices.
```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-seventeen
#| fig-cap: Probability tree for likelihood of sorting waste correctly, predicted by age and education.

knitr::include_graphics(here::here("data/04-model_data/sort_waste_tree.png"))

```

####  Likelihood of walking or cycling shorter distances
@fig-eighteen predicts the likelihood of individuals walking or cycling short distances based on their highest level of education and age. The root node identifies education level as the most significant factor. Individuals with postgraduate, undergraduate, or some university education are classified as “Already doing this or have done this” (n = 231, error = 51.9%). This indicates a strong relationship between higher education levels and the adoption of walking or cycling short distances.

For those with lower education levels (community/trade school, high school, or preferring not to answer), the model further splits based on age. The first division occurs at 65.5 years, distinguishing those “Very likely” to walk or cycle short distances (n = 15, error = 60%) from others. Additional splits by age at 33.5 years and 60.5 years, along with education, refine the classification. This stepwise approach demonstrates how age interacts with education to influence short-distance mobility choices.
```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-eighteen
#| fig-cap: Probability tree for likelihood of walking or cycling shorter distances, predicted by age and education.

knitr::include_graphics(here::here("data/04-model_data/short_distance_tree.png"))

```

## Model Estimation and Interpretation ##
The probability tree model estimates how the predictors age and education influence the likelihood of adopting climate-friendly behaviors. By splitting the data into branches based on these variables, the model helps identify decision rules that predict individuals' likelihood of taking actions such as reducing waste or investing in green products. We anticipate that younger individuals, due to greater environmental awareness, are more likely to engage in climate-positive actions, while those with higher levels of education may also be more inclined to adopt such behaviors, given their increased awareness of climate change and sustainable practices. The model estimates the parameters for age and education using regression techniques, providing coefficients that indicate how each factor influences the likelihood of taking climate-friendly actions.

# Results

## Model Summary Table
@fig-nineteen presents the accuracy of models predicting the likelihood of various climate-friendly actions based on demographic factors such as age and education. Each row represents a different climate action (e.g., home improvement, reducing hydro, minimizing car use), with the corresponding accuracy of the model's predictions. The accuracy values reflect how well the models, which use age and education as predictors, align with the actual data. An accuracy of 33 percent  for home improvement" indicates that the model correctly predicted the outcome 33% of the time for this action. These accuracy values help evaluate the effectiveness of age and education as predictors for each specific climate action.
```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-nineteen
#| fig-cap: This table presents the accuracy of models predicting the likelihood of various climate-friendly actions based on age and education. The accuracy values reflect how closely the model predictions align with the actual observed data. Higher accuracy values indicate better model performance.

# Load the model accuracy data from the saved CSV
model_accuracies <- read.csv(here::here("data/04-model_data/model_accuracies.csv"))

# Round numeric columns to whole numbers
model_accuracies[] <- lapply(model_accuracies, function(x) if(is.numeric(x)) round(x) else x)

# Check the first few rows of the data to ensure it loaded and rounded correctly
head(model_accuracies)

```

## Likelihood taking action by age
@fig-twenty shows the relationship between age and the likelihood of engaging in various climate-related actions. The x-axis represents the age groups, and the y-axis indicates the percentage of respondents reporting different likelihood levels, from "Already doing this" to "Very unlikely." The trend lines across each action show how engagement varies by age.

```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-twenty
#| fig-cap: Likelihood of Taking Climate Change Actions by Age Group. The stacked bar chart illustrates the percentage of respondents across six age groups (15-24, 25-34, 35-44, 45-54, 55-64, 65+) in terms of their self-reported likelihood to engage in nine distinct climate-related actions. Each color represents a different level of likelihood, from "Already doing this or have done this" to "Very unlikely." The black dashed trend line highlights how the likelihood of action changes across age groups, providing insights into generational differences in climate-related behavior.

knitr::include_graphics(here("data/03-figures_data", "likelihood_age_plot.png"))

```

## Likelihood taking action by education 
@fig-twenty-one displays how the likelihood of engaging in climate-friendly actions varies by educational attainment. The x-axis represents the proportion of respondents in each education category, while the y-axis lists the educational levels, from "Prefer not to answer" to "Postgrad." Each action is shown across separate facets, revealing how education level correlates with engagement in each behavior.Individuals with higher education levels generally show a higher likelihood of engaging in climate-friendly actions. This trend is evident across all behaviors, with respondents possessing undergraduate or postgraduate degrees more likely to adopt electric vehicles or purchase green products.

```{r}
#| warning: false
#| echo: false
#| message: false
#| label: fig-twenty-one

knitr::include_graphics(here("data/03-figures_data", "likelihood_education_plot.png"))

```

### Electric / Hybrid Vehicle Adoption
The model indicates that individuals with "Some Community/Trade School" or "Some University" education are the most likely to invest in electric or hybrid vehicles. This group is classified as "Very likely" (n = 47) with an error rate of 64%, suggesting that while many in this category are inclined to invest, there is notable variability. Among individuals with other educational backgrounds, age plays a key role. Younger individuals under 35 years old, particularly those classified as "Very likely" (n = 10, error = 30%), demonstrate a strong tendency to invest, reflecting high confidence in this prediction. However, another subset of younger individuals (n = 28) is "Somewhat likely," with a higher error rate of 61%, indicating less certainty. These trends suggest younger people are more open to adopting electric or hybrid vehicles, likely due to environmental awareness or familiarity with emerging technologies.

Older individuals, particularly those over 69 years old, are the least likely to invest. This group is classified as "Very unlikely" (n = 18) with an error rate of 61.1%, highlighting some uncertainty but an overall low likelihood of adoption. Middle-aged individuals (35 to 69 years old) show more mixed results, with some classified as "Somewhat unlikely" (n = 84, error = 54%) and others as "Somewhat likely" (n = 46, error = 57%). These findings suggest that barriers such as cost, resistance to change, or lower interest in alternative vehicles may deter older age groups.

In summary, the model suggests that younger individuals, particularly those under 35 years old, and those with mid-level education are the most likely to invest in electric or hybrid vehicles. In contrast, older individuals and those with minimal or advanced education are less likely to adopt these technologies. While the model provides insights into these trends, the error rates indicate variability in some predictions, particularly among "Somewhat likely" and "Somewhat unlikely" groups.

### Home Improvement for Energy Efficiency
The results reveal that age is the most significant predictor of climate-conscious home improvement behavior. Individuals aged 36 years or older are the most likely to already be taking or have taken such actions. This group represents the largest sample size (n = 223) but also exhibits a moderate error rate (59.6%), indicating some variation in their behaviors.

For younger individuals (under 36 years), educational attainment is a key determinant. Those with higher education levels, such as "Community/Trade School," "Some University," or "Postgraduate," are the most likely to take action ("Very likely") with an error rate of 55%. However, younger individuals with lower levels of education, such as "High School" or "Some Community/Trade School," are less likely to engage in these actions, being classified as "Somewhat likely" but with a higher error rate of 67%, indicating greater variability in this group.

Overall, individuals aged 36 or older are more likely to have already engaged in climate-conscious home improvements, regardless of education level. Younger individuals with higher education are more likely to adopt such changes in the future, while those with lower education are less likely, with predictions for this group being less certain. These findings underscore the importance of both age and education in shaping environmental engagement and suggest that outreach efforts might need to target younger, less-educated groups to increase participation.

### Consumption of Meat Alternatives
The model suggests that age is the strongest predictor of individuals’ likelihood to adopt meat alternatives for climate change mitigation. Younger individuals, particularly those under 34.5 years old, show varying tendencies. Among this group, those aged 31.5 years or younger are divided into “Somewhat likely” (n = 25, error = 44%) and “Somewhat unlikely” (n = 62, error = 62.9%), indicating considerable variability in predictions. These findings suggest that while younger individuals are more open to meat alternatives, their likelihood of adoption is not uniform.

For individuals aged 34.5 years or older, education becomes a significant predictor. Those with community/trade school, undergraduate, or postgraduate education are more likely to consider or adopt meat alternatives. Within this group, further splits by age highlight nuanced behaviors. For instance, individuals over 65.5 years old are classified as “Already doing this or have done this” (n = 30, error = 56.7%), reflecting a subset of older adults who actively embrace meat alternatives, possibly due to increased awareness or health considerations.

Conversely, individuals with high school education in this older age group show lower likelihoods of adopting meat alternatives, with many classified as “Somewhat unlikely” (n = 22, error = 68.2%). These results underscore the interplay of age and education, where younger individuals and those with higher education levels are more inclined toward adopting meat alternatives. However, the variability in error rates across groups indicates that other unmeasured factors, such as income or cultural attitudes, may also influence these decisions.

### Minimizing Car Use
The decision tree indicates that the majority of individuals in the dataset are classified as "Already doing this or have done this," with a sample size of 320 and an error rate of 61.9%. This classification implies that minimizing car use is a prevalent behavior among the population studied. However, the relatively high error rate suggests notable variability within this group, indicating that while many individuals report minimizing car use, there may be discrepancies or inconsistencies in the data.

The lack of further splits in the decision tree highlights the uniformity of responses or the insufficiency of the input variables to explain differences in behavior. This uniform classification could reflect widespread adoption of car use minimization practices due to shared motivations, such as environmental awareness or accessibility to alternatives like public transportation or carpooling.

In summary, the model reveals that minimizing car use is a broadly practiced behavior within the sample population. However, the high error rate points to potential data quality issues or the influence of unmeasured factors. Future research could benefit from incorporating additional variables, such as geographic location, income level, or access to transportation infrastructure, to better understand the drivers and barriers to minimizing car use.

### Purchasing Green Products
The model reveals distinct patterns regarding the likelihood of purchasing green products based on age and education. Among individuals aged 37.5 years or older, a notable segment is classified as "Somewhat likely" (n = 66, error = 54.5%). However, older individuals with age thresholds of 47.5 years and above show increased tendencies to fall into the category of "Already doing this or have done this." For instance, individuals aged over 50.5 years are predominantly classified as "Already doing this or have done this" (n = 7, error = 42.9%). These findings suggest that age plays a significant role in influencing behavior, with older age groups showing stronger commitment to green product purchases.

For younger individuals under 37.5 years, the likelihood of purchasing green products varies based on education and finer age splits. Those under 27.5 years are more likely to be classified as "Very likely" (n = 24, error = 54.2%), while individuals under 22.5 years exhibit an even higher likelihood, with a smaller error rate (n = 8, error = 37.5%). This trend indicates that younger individuals, particularly those in the youngest age brackets, are more inclined toward green product purchasing, possibly reflecting heightened environmental awareness or greater receptiveness to sustainable practices.

Overall, the decision tree highlights the interplay of age and education in predicting green product purchasing behavior. Younger individuals, especially those with mid-range education, demonstrate a stronger inclination toward adopting sustainable practices. In contrast, older individuals, particularly those above 50 years, are more likely to have already adopted these behaviors. While the model provides valuable insights, the variability in error rates across different nodes suggests that other unmeasured factors may also influence green product purchasing behavior.

### Reducing Hydro Usage
The findings from this decision tree analysis suggest that age and educational attainment are pivotal factors shaping individuals' likelihood of reducing hydro usage. Younger individuals, particularly those under 34.5 years, exhibit greater openness and enthusiasm for hydro-reduction practices. This trend likely reflects their increased exposure to sustainability discourse, progressive values, and familiarity with energy-saving technologies. Similarly, older individuals (above 64.5 years) might also demonstrate a willingness to reduce hydro usage, potentially driven by long-term environmental awareness or economic considerations during retirement. In contrast, middle-aged individuals show a more nuanced relationship with hydro-use reduction, as their decisions are influenced by their educational background, with those holding higher educational qualifications demonstrating greater likelihood to act.

These results carry significant implications for policy and program design aimed at reducing hydro consumption. Younger individuals can serve as early adopters and advocates for hydro-saving technologies, making them a critical demographic for energy campaigns and innovative solutions. For older individuals, policymakers can emphasize the economic benefits of hydro-reduction initiatives, such as savings on utility bills, to motivate action. Addressing the middle-aged demographic requires more targeted approaches that consider educational disparities. Educational interventions, including workshops and informational campaigns, could bridge gaps for those with lower education levels, helping to increase their likelihood of engagement. Overall, these insights highlight the need for age- and education-specific strategies to effectively promote widespread hydro-use reduction.

### Reducing Waste
The results show that younger individuals under 34.5 years are more likely to be categorized as "Somewhat likely" or "Very likely" to reduce waste, particularly when they have higher levels of education. This suggests that environmental awareness among younger generations is strongly linked to educational attainment. In contrast, older individuals over 75.5 years are predominantly categorized as "Already doing this or have done this," reflecting long-standing habits and greater experience with resource conservation.

For individuals aged 34.5 to 75.5 years, the likelihood of reducing waste varies based on both age and education. Those with higher educational attainment are more likely to have already adopted waste-reduction behaviors, while those with lower education levels are less consistent in their engagement. The errors associated with the predictions indicate variability within these groups, emphasizing the need for targeted interventions to improve participation rates.

The findings suggest clear implications for policy and outreach. For younger individuals, campaigns focusing on leveraging their education and enthusiasm for sustainability can encourage even higher engagement. For middle-aged groups, addressing educational gaps through workshops or accessible resources can boost participation. For older individuals, reinforcing their existing waste-reduction practices with supportive policies, such as incentives for recycling or composting, can sustain their efforts. Tailoring strategies to these specific groups ensures a more effective approach to fostering waste reduction.

### Sorting Waste Correctly
The results indicate that individuals with higher educational attainment (Community/Trade School, Undergraduate, or above) overwhelmingly report "Already doing this or have done this" when it comes to sorting waste. This suggests that education enhances awareness and engagement with sustainable practices. However, the moderate error rate (37.3%) implies variability within this group, potentially due to differences in access to infrastructure or levels of personal commitment.

Among those with a High School education, age emerges as a crucial factor. Younger individuals under 37 years are also likely to have sorted waste, though with higher inconsistencies. Older individuals in this category exhibit a stronger likelihood of intending to sort waste ("Very likely"), reflecting a growing awareness or changing attitudes with age. These findings highlight the nuanced relationship between age, education, and environmental behavior.

The implications of this model suggest the need for tailored interventions. For individuals with lower educational attainment, targeted educational campaigns or accessible resources can bridge knowledge gaps and improve waste-sorting habits. For younger groups, embedding sustainability practices into school curricula can reinforce early habits. For older populations, providing convenient infrastructure or incentives for sorting waste could further encourage their commitment. Overall, the results emphasize the importance of addressing both educational and age-specific barriers to enhance waste-sorting behaviors.

### Walking / Cycling Short Distances
The model highlights a clear link between education and walking or cycling short distances. Individuals with higher education levels are more likely to engage in these activities, suggesting that education fosters awareness or access to environments conducive to active transportation. However, the moderate error rate in this group (51.9%) implies that other unmeasured factors, such as urban infrastructure or cultural attitudes, may also influence the behavior.

Age emerges as a secondary but critical determinant, particularly for individuals with lower education levels. Those aged over 65.5 years are classified as “Very likely” to adopt walking or cycling, which may reflect a shift toward health-conscious behaviors among older adults. In contrast, individuals under 33.5 years show varying likelihoods, with subgroups classified as “Somewhat likely” or “Already doing this,” reflecting the nuanced impact of life stage and environmental factors.

These findings suggest that promoting active transportation should consider demographic factors, such as age and education, alongside broader interventions like improved urban design or targeted awareness campaigns.

## Overall Summary of Results:
The results indicate that both age and education significantly influence the likelihood of engaging in climate-related actions. Younger respondents tend to be more likely to adopt behaviors such as consuming meat alternatives and minimizing car use, while older individuals are more inclined to engage in actions that require larger investments or long-term commitments, such as reducing hydro usage and investing in energy-efficient home improvements. Higher educational attainment is also strongly linked to a greater likelihood of engaging in climate-friendly behaviors, particularly those that involve financial investment, like adopting electric vehicles and consuming meat alternatives. However, simpler actions such as sorting waste correctly are more widely reported across all education levels. These generational and educational differences highlight the importance of factors such as income, lifestyle, and environmental awareness in shaping individuals' climate-related behaviors. In light of these findings, it is crucial to consider demographic factors when designing policies or campaigns aimed at encouraging climate action, as different age and education groups exhibit varied levels of engagement with both simple and more complex sustainable practices.

# Discussion

## Age and Climate-Friendly Behaviors {#sec-first-point}
The findings indicate that younger individuals are more likely to engage in climate-friendly actions perceived as low-cost and easier to implement, such as consuming meat alternatives and reducing car usage. The model results show that these behaviors have relatively higher predicted likelihoods among younger participants, aligning with existing research on generational differences in environmental attitudes. This trend may be driven by heightened environmental awareness and increased exposure to climate change discussions among younger generations, often through social media and educational initiatives. These results suggest that lifestyle factors and generational attitudes significantly shape behavioral patterns, with younger individuals gravitating toward actions that are both socially encouraged and accessible.

## The Role of Education in Sustainable Actions {#sec-second-point}
Education significantly influences individuals' likelihood of engaging in more substantial climate-friendly behaviors. Our models demonstrate that individuals with higher educational attainment are more likely to adopt behaviors requiring greater financial or logistical investment, such as purchasing electric vehicles or making energy-efficient home improvements. The accuracy of the models predicting these behaviors is relatively lower than for simpler actions, suggesting that education is an important but not sole determinant. This relationship between education and sustainability underscores the importance of knowledge and informed decision-making in supporting climate-friendly actions, particularly those with long-term environmental benefits. The data suggests that higher education increases both environmental awareness and the capacity to implement financially demanding sustainable practices, which are often inaccessible to those with lower educational levels.

## The Influence of Climate Change Knowledge {#sec-third-point}
Self-reported knowledge of climate change also plays a role in the likelihood of adopting sustainable behaviors. Our analysis shows that individuals with greater awareness are more proactive in making environmentally conscious choices, particularly for actions like reducing waste and conserving energy. However, the model accuracy for some knowledge-based behaviors remains moderate, highlighting a gap between awareness and action. This suggests that while raising awareness is critical, it must be paired with policies that reduce structural barriers to adopting more expensive or long-term sustainable actions. For example, financial incentives or subsidies for purchasing energy-efficient appliances could bridge this gap, ensuring individuals are both informed and able to make practical changes.

## Limitations

One significant limitation of this study is its reliance on self-reported survey data, which is inherently prone to biases such as social desirability bias and recall errors. Respondents may overestimate their environmental knowledge or exaggerate their participation in climate-friendly behaviors to present themselves in a favorable light. This introduces uncertainty into the accuracy of the models, as the input data may not fully reflect actual behaviors or knowledge levels. Additionally, self-reported data often lacks the granularity needed to explore complex interactions between socio-economic factors, such as income and household size, which can significantly influence sustainable actions.

Another limitation arises from the use of summary data from the 2018 and 2021 surveys instead of individual-level data. While the summary statistics provided valuable insights into broad demographic trends, they limited the study’s ability to explore more nuanced patterns and interactions. For example, the summary data showed differences in education and age trends between 2018 and 2021, but without individual-level data, it was difficult to determine the direct effect of socio-economic factors on climate action likelihood. Future studies with access to raw, individual-level data could offer more precise estimates and a deeper understanding of these dynamics.

## Challenges with Survey Data
Survey data presents additional challenges, such as non-response bias, where certain demographic groups may be underrepresented, leading to skewed results. Additionally, the framing of survey questions can influence how respondents perceive and answer them, introducing further variability in the data. The inability to verify self-reported behaviors also limits the reliability of the conclusions drawn from such datasets. Despite these challenges, surveys remain a valuable tool for capturing perceptions and self-reported behaviors, but their findings should be interpreted with caution and, when possible, supplemented with objective data sources.

## Future Directions
Future research should aim to identify the specific barriers preventing certain demographic groups from adopting climate-friendly behaviors. This could involve examining factors such as income, access to resources, and perceived obstacles to sustainable practices. Additionally, targeted interventions, such as financial subsidies, public awareness campaigns, or infrastructure improvements, could make sustainable actions more accessible and affordable. Longitudinal studies would be particularly valuable for tracking how awareness of climate change influences sustained behavioral changes over time. Comparing individual-level data across different years could also provide insights into the long-term impact of educational and policy interventions on climate-friendly behaviors.

Lastly, integrating objective data, such as energy consumption records or transportation usage statistics, could validate and enhance the accuracy of survey-based findings. This combined approach would offer a more comprehensive understanding of how demographic and socio-economic factors influence the adoption of sustainable actions, ultimately informing more effective climate policies and interventions.

# Conclusion {#sec-conclusion}
This study provides valuable insights into the demographic factors that influence individuals' likelihood to engage in climate-friendly behaviors, particularly age, education, and climate change knowledge. Younger individuals and those with higher levels of education are more likely to adopt sustainable behaviors, particularly those that require greater financial investment or long-term commitment. However, barriers such as financial constraints and access to resources remain significant challenges. Addressing these barriers through targeted policies and interventions could help make sustainable behaviors more accessible and the norm across all demographic groups. Future research should explore the underlying reasons for resistance to adopting climate-friendly behaviors and how external factors can help overcome these challenges to achieve widespread climate action.

\newpage

\appendix

# Appendix {#sec-appendix}

# Idealized methodology {#sec-idealized-meth}
## Survey objectives
The primary objective of this study is to explore the factors influencing the likelihood of individuals engaging in climate-friendly behaviors. Specifically, the study aims to identify the age groups that are least likely to participate in sustainable actions and investigate how education levels impact climate-friendly behaviors. The research also seeks to determine whether individuals who report being relatively informed about climate change causes are more likely to take action to mitigate it. The underlying assumption is that individuals with greater knowledge of climate change should be more inclined to adopt behaviors that reduce its impact. Additionally, the study will examine the barriers that prevent people from engaging in sustainable behaviors, with the goal of identifying strategies to make these actions more accessible, affordable, and easy to adopt. The overarching aim is to develop solutions to increase the likelihood of participation in climate-friendly behaviors by removing existing barriers and promoting sustainability across different demographic groups. Part of this involves determining the most effective platforms for disseminating information about climate change, specifically by identifying which communication methods resonate most with different age groups. The research will also explore the role of education systems in fostering climate awareness and encouraging sustainable actions from a young age. Ultimately, the findings aim to provide actionable solutions to promote wider participation in climate-friendly actions, contributing to the fight against climate change.

## Sampling approach
This study will target a diverse set of respondents across various age groups and demographic characteristics to gain a comprehensive understanding of climate-friendly behaviors and their drivers. The sample will be stratified to ensure representation across youth and adult demographics, enabling a comparison of climate knowledge and behaviors between generations. The study will start by sampling youth as young as age 12. This age group is chosen because children are old enough to understand the basics of climate change and the actions required to address it but still impressionable enough for their behaviors to be influenced by education and social interventions. By including participants starting at age 12, we aim to assess how effective climate change education in schools is and whether it is shaping climate-friendly behaviors in the younger generation.

In addition to youth, adults aged 18 and older will also be included to provide insight into longer-term behavioral trends, which may have been shaped by earlier education or a lack of education on the subject. The sampling will be stratified according to age, education level, income, and geographic location to capture a broad spectrum of experiences and perspectives on climate change and sustainability. Age groups will be divided into the following ranges: 12-17, 18-24, 25-34, 35-44, 45-54, 55-64, and 65+, while education levels will range from middle school to postgraduate education. Income will be categorized as low (<$30,000), middle ($30,000–$99,999), and high (>$100,000), and participants will also be grouped based on urban, suburban, or rural geographic locations. This stratified sampling approach ensures a diverse representation of individuals, allowing for meaningful comparisons of climate knowledge and behaviors across different demographic groups.

## Respondant recruitment 
Participants will be recruited using a multi-channel approach to ensure broad representation across various age groups and geographic locations. For youth aged 12-17, recruitment will occur through schools, with cooperation from middle and high schools. District educational boards or individual teachers will be approached to distribute information about the survey, and parental consent will be obtained before participation. To recruit adults, online surveys will be distributed via social media, email lists, and relevant climate-focused online communities. Additionally, outreach will be conducted in community centers and public spaces to capture respondents from rural or underserved areas. Flyers, local advertisements, and community events will help ensure that the sample is representative of both urban and rural populations.

Incentives such as gift cards or a donation to a climate-focused charity will be offered to participants to encourage engagement and improve response rates, especially among younger participants who may be less inclined to participate without an incentive.

## Data Validation 
To ensure the validity and reliability of the data, several validation mechanisms will be employed. Responses will be reviewed for internal consistency, such as cross-checking whether respondents who claim high levels of climate change knowledge also report engaging in corresponding climate-friendly behaviors. Participants who provide inconsistent answers or fail to complete the survey will be excluded from the analysis. Additionally, respondents may be contacted for clarification if their answers are ambiguous or incomplete. This will ensure that the data accurately reflects the participant's true attitudes and behaviors.

# Weighting and Data adjustments
The data collected will be weighted to account for any imbalances in demographic representation within the sample. Statistical weighting will adjust for factors such as age, education level, income, and geographic location to ensure the sample accurately reflects the broader population. Weighting will correct for any over- or under-representation of specific demographic groups, allowing for more generalizable findings. Additionally, data will be adjusted for non-response bias, with greater weight given to underrepresented groups (e.g., certain age groups or income levels) to correct for gaps in participation.

## Budget
The budget for this study will cover several key areas essential for effective data collection and analysis. A portion of the budget will be allocated to the survey platform and software tools needed for data collection, such as online survey tools and platforms for both youth and adult respondents. Since incentives are critical to encouraging participation, especially from younger individuals, funds will also be used to provide gift cards or donations to climate-related charities as incentives. To reach a diverse demographic, outreach costs will also be factored in, including the production and distribution of flyers, local advertisements, and outreach at community centers or schools. Lastly, resources will be dedicated to data analysis, including the purchase of statistical software like SPSS or R for cleaning, coding, and analyzing the responses. The final portion of the budget will cover any miscellaneous costs such as obtaining parental consent forms, printing, and mailing physical surveys or communication materials.

## Survey design
The survey design will be structured to capture a range of demographic and behavioral data in order to address the research objectives. The survey will begin with demographic questions to capture essential background information, including age, education level, income, and geographic location. These variables will allow the analysis to identify how climate-friendly behaviors and knowledge vary across different groups. The primary focus will be on participants' climate change knowledge, specifically assessing their awareness of its causes, impacts, and potential solutions. Respondents will be asked to rate their perceived level of knowledge about climate change on a Likert scale, ranging from "not informed" to "very informed." They will also be prompted to identify key contributors to climate change, as well as actions individuals can take to mitigate its effects.

The next section of the survey will focus on climate-friendly behaviors, asking respondents to self-report on their engagement in specific actions such as reducing car use, purchasing electric vehicles, eating plant-based meals, recycling, and reducing household energy consumption. These behaviors will help determine the extent to which knowledge of climate change is translating into actions, as well as highlight potential gaps in engagement.

Following this, the survey will include a section on barriers to sustainable behaviors, exploring factors that may be preventing individuals from adopting more climate-friendly actions. Participants will be asked to identify reasons such as cost, convenience, lack of knowledge, or skepticism. This will allow the study to pinpoint the most significant obstacles to widespread adoption of sustainable practices and provide insight into potential solutions.

The final section will focus on how respondents prefer to receive information about climate change. Given the varying effectiveness of different communication channels for different age groups, the survey will include questions on the most effective platforms for disseminating climate change education, such as social media, school programs, TV ads, and email. By identifying these preferences, the study will be able to recommend the most suitable methods for spreading climate change awareness tailored to specific demographics.

Each section of the survey will include a combination of multiple-choice, Likert scale, and open-ended questions to capture both quantitative and qualitative data. The open-ended responses will allow for more nuanced insights into the reasons behind individuals' climate-related behaviors and barriers. The design will aim to balance simplicity with comprehensiveness to ensure that the survey is engaging and easy to complete while still gathering sufficient data to answer the research questions effectively.

## Tradeoffs and limitations
While this methodology is designed to be comprehensive, there are inherent trade-offs and limitations that must be considered. One key limitation is the reliance on self-reported data. Since participants will be answering questions based on their own recollections and perceptions, there is the potential for bias, such as over-reporting sustainable behaviors or under-reporting barriers to climate action. This could skew the findings, particularly if individuals respond in a socially desirable manner. Additionally, while efforts will be made to capture a representative sample, there is a possibility of sampling bias, especially in rural areas or among specific income groups who may have limited access to the survey. This could affect the generalizability of the results. Further, survey fatigue is a potential issue, particularly with younger participants who may lose focus or fail to complete longer surveys. To mitigate this, the survey will be kept concise, though the breadth of questions may still pose challenges. Finally, while statistical weighting and adjustments for non-response bias will be implemented, there is always some degree of uncertainty regarding the effectiveness of these methods in fully correcting for bias or imbalances in the sample. These limitations must be taken into account when interpreting the findings of this study.

## Idealized survey questions
Thank you for your participation in the 2024 Climate Change Survey. This survey aims to gather information about public attitudes and behaviors related to climate change, with a focus on understanding the factors that influence people's engagement in climate-friendly actions. Your participation is entirely voluntary, and you may withdraw at any time, for any reason, with no questions asked.

This survey collects data regarding your awareness of climate change, the actions you take to mitigate it, and the barriers that may prevent you from engaging in more sustainable behaviors. The data you provide will be kept confidential and will be used solely for research purposes. This survey is anonymous, and your responses will not be traceable back to you. The goal of this survey is to better understand the motivations, challenges, and opportunities related to climate action, with a view to improving strategies for promoting sustainability.

If you have any questions or concerns about this survey or its methodology, please feel free to contact Lexi Knight via email at lexi.knight\@mail.utoronto.ca. Any correspondence will remain confidential and will not be shared with any external parties.

**Screening and Consent:**

By checking this box, I consent to this survey collecting information about my awareness of climate change, the actions I take to address it, and the factors that influence my behavior for research

Correspondence will not be shared with any external parties.
-    I consent.

**Demographic Information: **
1. Whats your age? 12-17, 18-24, 25-34, 35-44, 45-54, 55-64, 65+, Prefer not to say
2. What is the highest level of education that you've completed? Some high school, High school diploma, Diploma / post secondary certificate (college, trade school). Bachelor's degree, Master's degree, Doctorate degree.

**Climate Change knowledge and Awareness:**
3. How would you rate your knowledge of the causes of climate change? Very informed, Somewhat informed, Neutral, Somewhat Uninformed, Very uninformed.
4. How confident are you in your understanding of how to mitigate climate change through individual actions? Very confident, Confident, Neutral, Somewhat confident, Not confident.

**Climate-Friendly Action: **
5. Which of the following climate-friendly actions do you regularly engage in? (Select all that apply). Reduce personal vehicle use (e.g., carpooling, using public transport, biking or walking shorter distances), Use / install energy-efficient appliances / devices (e.g., LED bulbs, smart thermostats), Eat a plant-based diet or reduce meat consumption, Recycle and compost, Purchase eco-friendly products (e.g., sustainable clothing, eco-conscious brands), Use renewable energy sources (e.g., solar panels, wind energy), None of the above, Other.
6. How often do you consider the environmental impact of your purchases? Always, Often, Sometimes, Rarely, Never

**Barriers to Sustainable Actions: **
7. What factors prevent you from engaging in more climate-friendly behaviors? (Select all that apply) Cost of sustainable products or services, Lack of time, Lack of knowledge about how to make sustainable choices, Convenience (e.g., environmentally harmful options are more accessible), Skepticism about the effectiveness of individual actions, No barriers (I already engage in sustainable behaviors), Other (please specify)
8. Do you feel that climate-friendly behaviors are affordable for people in your community? Yes, No, Not sure
9. Which of the following reasons best describe why you do not participate in more climate-friendly actions? (Select all that apply) I do not believe my individual actions will make a difference, I find it too difficult to make sustainable choices in my daily life, It is too expensive to adopt sustainable behaviors, I don’t know where to start or how to make a meaningful impact, Sustainable products or services are not easily available in my area, I do not feel that climate change directly affects me or my community, I am unsure about what actions are most effective in reducing climate change, I am not sure how to balance sustainable actions with my current lifestyle
10. If you have avoided taking climate-friendly actions, what do you believe would make it easier for you to participate in these behaviors? (Select up to three) Lower cost of sustainable products or services, More convenient access to sustainable options, Clearer information on how individual actions can make a difference, More education or awareness about climate change, Incentives (e.g., government subsidies, rewards) to take action, More social pressure or norms encouraging sustainable behavior, Other (please specify)
11. How much do you trust the information provided by the following groups on climate change? Government, Environmental NGOs, Local community organizations, Media (TV, online news), Social media influencers/bloggers, Scientists, I do not trust any of these sources
12. If the government provided more support for individuals to take climate-friendly actions, would this lead you to taking more environmentally friendly actions (e.g., incentives, education programs, reduced cost)? Yes, No, Not sure
13. How strongly do you agree with the following statement: "Taking climate-friendly actions is a personal responsibility." Strongly disagree, Disagree, Neutral, Agree, Strongly agree
14. How strongly do you agree with the following statement: "The government or businesses should take more responsibility for addressing climate change than individuals." Strongly disagree, Disagree, Neutral, Agree, Strongly agree
15. Do you think the reluctance to take action against climate change more due to a lack of knowledge about how to get involved, or is it primarily because sustainable actions are inconvenient for people? Lack of knowledge about how to take action, Inconvenience of sustainable actions, Both, Neither, Unsure

**Education and Information: **
16. How effective do you think current school programs are in educating students about climate change? Very ineffective, Somewhat ineffective, Neutral, Somewhat effective, Very effective
17. Which platforms do you find most effective for receiving information about climate change? Social media (e.g., Instgram, Facebook, Twitter, TikTok), School programs, TV/Youtube, News articles, Websites, Podcasts, Community events, Email newsletters.

**General Perception: **
18. Do you believe that individual actions can make a significant difference in combating climate change? Yes, No, Not sure

19. In your opinion, which of the following should be prioritized to encourage more sustainable behaviors? (Select up to two), Making sustainable actions more affordable, Providing more education and awareness about climate change, Improving the accessibility of sustainable options, Encouraging government policies and incentives for sustainable behaviors, Creating a societal norm where sustainable actions are expected and practiced

**Confirmation Message**

Thank you for your response. We greatly appreciate the time, effort, and honesty you dedicated to completing this survey. Your answers have been successfully recorded and will contribute significantly to our research!

# Additional data details
After data collection, the responses will be cleaned, coded, and analyzed using appropriate statistical methods. Analysis will focus on identifying correlations between demographic factors (age, education, income, geography) and engagement in climate-friendly behaviors. The data will also be examined for patterns in barriers to participation and preferences for information delivery. These findings will provide the basis for actionable recommendations to improve climate change education and encourage greater participation in sustainable behaviors.

## Simlation Process {sec-simulation-process}
For the purposes of this study, I simulated data to investigate various factors related to climate change behaviors, including demographics, education levels, awareness of climate change causes, and preferences for communication methods. The simulated dataset consists of 404 entries, each representing an individual. To create this dataset, I used the `tidyverse` and `arrow` packages in R and set a seed value `(set.seed(853))` to ensure the results were reproducible. I generated the following variables:

The age variable was sampled randomly from a range between 18 and 100 years old. For education, I created several categories, such as "high school or less," "some community college/trade school," and "postgraduate/professional school," with an option for "prefer not to answer." The informed variable was categorized into four levels, from "extremely informed" to "not at all informed," reflecting the extent to which individuals felt knowledgeable about climate change causes. To capture the likelihood of engaging in climate-friendly actions, I simulated the likelihood of taking action variable, with responses ranging from "already doing this or have done this" to "very unlikely." Finally, the communication method variable explored how individuals preferred to receive climate-related information, with options such as "Toronto.ca website," "social media platforms," and "advertising campaigns," among others.

I used random sampling to generate values for each of these variables, ensuring a broad range of responses. After generating the data, I summarized it to verify its structure and consistency. The dataset was then saved as a `.parquet` file for further analysis. This approach allowed me to simulate a diverse set of responses and explore potential relationships between demographic factors, climate awareness, and willingness to engage in climate-friendly behaviors, all while maintaining the flexibility of synthetic data.

## Data cleaning {#sec-data-cleaning}

### 2018 Individual Data
The cleaning process for the 2018 individual data began with selecting six key questions from the survey, focusing on variables that would provide insights into individual perceptions and behaviors regarding climate change. The corresponding columns were extracted and renamed with meaningful, descriptive names for clarity. The selected variables included age, the extent to which individuals consider themselves informed about the causes of climate change, the likelihood of taking specific climate actions, reasons for inaction (if they indicated they were unlikely to act), the highest level of education completed, and preferred methods for the city to deliver information about climate change and climate action.

After selecting and renaming the columns, minor formatting inconsistencies were addressed. For instance, a missing space in the "verylikely" response category was corrected to "very likely." Additionally, certain values were reformatted to ensure compatibility with visualization tools, making the dataset easier to plot later. Unlike other datasets that often contain missing values or duplicate entries, this dataset was notably clean. There were no missing responses, and all questions were answered, allowing for a comprehensive analysis of survey participants' perceptions and behaviors. The dataset offered valuable insights into individuals’ likelihood of engaging in climate actions and their preferred communication channels for climate information.

Throughout the cleaning process, no observations were removed, as every response was preserved to accurately represent the survey data. Despite encountering a few incorrect variable formats, these were easily corrected without data loss. The dataset provided a rare opportunity to analyze a complete and consistent set of individual responses, enabling a thorough exploration of both demographic characteristics and personal perceptions related to climate change.\

### 2018 Summary Data
Cleaning Process for the 2018 Summary Data
To create the 2018 summary data, I built off the cleaned 2018 individual data. This approach was necessary because the 2021 data was only available in summarized form, not at the individual level. By summarizing the 2018 data in a comparable format, I could facilitate a meaningful comparison between the two years. The first step involved loading the cleaned 2018 individual dataset and mimicking the structure of the 2021 summary data as closely as possible.

For age, I created the same age categories used in the 2021 dataset to ensure consistency between the two datasets. I then calculated the percentage of individuals within each age group and created a table summarizing this information. Similarly, I summarized the highest level of education completed by respondents, presenting it as a percentage of the total. However, a slight inconsistency emerged during this process, as the education levels and their descriptions differed between 2018 and 2021. Despite this discrepancy, I aligned the categories as closely as possible to maintain comparability.

Next, I calculated the percentage of respondents who reported feeling informed about the causes of climate change and summarized the likelihood of taking various climate actions. This step was somewhat more complex, as the dataset contained nine different actions, each rated on a five-point scale of likelihood. I created a table showing the percentage likelihood of taking each specific action. Additionally, I analyzed the reasons respondents provided for being unlikely to take certain actions, which were only collected if they had indicated “unlikely” in the previous question. These reasons were summarized as counts rather than percentages, reflecting the conditional nature of the responses.

Finally, I summarized the preferred methods for receiving climate-related information from the city, presenting this data as a percentage. This comprehensive summarization of the 2018 individual data provided a consistent and comparable dataset for analyzing trends and changes in perceptions and behaviors between 2018 and 2021.

### 2021 Summary Data 
The cleaning process for the 2021 summary data focuses on ensuring consistency and accuracy for six specific variables: age, education, extent informed, likelihood to take action, and reasons for not taking action. I first extracted data from the relevant sheets and columns in the 2021 raw dataset using read_excel from the tidyverse. For each category, I selected specific rows and converted raw data into percentages. The age and education data were cleaned and saved into separate data frames (age_summary_21 and education_summary_21), while other categories like "extent informed" and "likelihood to take action" required cleaning across multiple columns, with data transformation into percentages. Missing values and outliers were handled by filtering invalid entries and ensuring all values were consistent.

Instead of merging datasets for 2018 and 2021, I aligned the 2018 summary data to match the structure and categories of the 2021 data to facilitate comparison. This alignment was necessary because the categories and structure of the two datasets differed, and I modified the 2018 dataset to resemble the 2021 format as closely as possible, allowing for a more direct comparison. By focusing on ensuring that categories and percentages were comparable between years, I preserved the integrity of both datasets.

Finally, I saved each cleaned dataset into formatted tables using tinytable for LaTeX compatibility. These cleaned datasets were essential for analysis, ensuring that all variables were represented as percentages without any missing or erroneous data. This process also ensured that the dataset was ready for comparison with the 2018 data, with clear, consistent categories across both years.

\newpage

# References